---
title: "AlcGen Pupil Data Analysis"
author: "Zihan Bai"
date: "`r Sys.Date()`"
description: "This script analyze pupil data for the AlcGen study. It reads preprocesed data, performs cleaning and analysis, and outputs figures"
output: html_document
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("pacman)
library(pacman)
p_load(readr, dplyr, ggplot2, PupillometryR, PupilPre, shiny, tidyverse, stringr, emmeans, sjPlot, nlme, patchwork)


# Change this to whatever you are working on
subject <- c("L003", "L004", "L006", "L007", "L009")  # Replace with actual subject
task <- list(c("conditioning")) # Replace with actual task (conditioning/generalization/csmem/usmem)
task_abbr <- list(c("cond")) # abbreviation of the task for naming the variable (cond/gen/csmem/usmem)
phase <- list(c("cs_onset")) # Replace with actual phase (cs_onset/anticipation/us_onset/gs_onset)

# Define paths
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"
output_path <- file.path(base_path, "eye", "results")

input_path_eye <- file.path(base_path, "eye", "preprocessed", subject, sprintf("%s_%s_%s_preprocessed.csv", subject, task, phase))
input_path_beh <- file.path(base_path, subject)

```

```{r}

all_subjects_cond <- list()
all_subjects_cond[[subject]] <- cond_us_cleaned

process_subject <- function(subject) {
  # Define paths
  input_path_eye <- file.path(base_path, "eye", "preprocessed", subject, sprintf("%s_%s_%s_preprocessed.csv", subject, task, phase))
  input_path_beh <- file.path(base_path, subject)
  
  # Load and process behavior
  process_beh_cond(subject, task, phase, task_abbr)
  
  # Load and process eyes
  
}
```


# Initial function to load behavioral data - call it later

```{r}
process_beh_cond <- function(subject, task, phase, task_abbr) {
    beh_pattern <- sprintf("%s_task_%s_events_.*\\.csv$", subject, task)
    beh_files <- list.files(input_path_beh, pattern = beh_pattern, full.names = TRUE)
    
    if (length(beh_files) != 1) {
      stop("Either no or multiple behavioral data files found. Please check the directory and pattern.")
      }
    beh_raw <- read_csv(beh_files[1], show_col_types = FALSE)
  
    # Assign the data to a variable with customized name
    assign(sprintf("beh_%s_raw", task_abbr), beh_raw, envir = .GlobalEnv)
    
    beh_cond <- beh_cond_raw %>%
      mutate(
      stim = case_when(
        category == "neut" ~ "neut",
        category == "alc" ~ "alc",
        category == "null" ~ "null",
        category == "null_alc_surprise" ~ "null",
        category == "null_neut_surprise" ~ "null",
        category == "alc_null_surprise" ~ "alc",
        category == "neut_null_surprise" ~ "neut"),
      incong_loc = case_when(
        category %in% c ("alc_null_surprise", "null_alc_surprise", 
                      "null_neut_surprise", "neut_null_surprise") ~ TRUE,
        TRUE ~ incong_loc))
    
    vcs_list <- c(unique(beh_cond[grepl('neut', beh_cond$stim),]$cs),
              unique(beh_cond[grepl('alc', beh_cond$stim),]$cs),
              unique(beh_cond[grepl('null', beh_cond$stim),]$cs))
    vcs_ord <- c('neut', 'alc', 'null')

    # Find the number of exposure ppl have seen for each CS type
    beh_cond <- beh_cond %>%
      mutate(
        cs = case_when(
          cs %in% vcs_list[1] ~ vcs_ord[1],
          cs %in% vcs_list[2] ~ vcs_ord[2],
          cs %in% vcs_list[3] ~ vcs_ord[3],
          TRUE ~ cs)) %>%
      arrange(trial) %>%
      group_by(cs) %>%
      mutate(iter = row_number()) %>%
      ungroup()
}
```

# Conditioning Task

## Recode behavior

```{r}

```

```{r}
# # What is the congruent ROI for each CS type for this subject?
# cong_ROI <- beh_cond %>%
#   select(cs, incong_loc, pos_item) %>%
#   filter(cs != "null" & incong_loc == "FALSE") %>%
#   group_by(cs) %>%
#   summarize(cong_ROI = unique(pos_item)) %>%
#   ungroup() %>%
#   print()
```

## Recode eye

```{r}
# Load the preprocessed eye-tracking data
eye_cond_cs_raw <- read_csv(input_path_eye, show_col_types = FALSE)

eye_cond_cs <- eye_cond_cs_raw %>% 
  select(subject = Subject,
         trial = TRIAL_INDEX,
         time = Time,
         pupil = Pupil,
         cs = cs,
         ia = RIGHT_INTEREST_AREA_ID,
         gazex = Gaze_X,
         gazey = Gaze_Y) %>%
  mutate(
    cs = case_when(
      cs %in% vcs_list[1] ~ vcs_ord[1],
      cs %in% vcs_list[2] ~ vcs_ord[2],
      cs %in% vcs_list[3] ~ vcs_ord[3],
      TRUE ~ cs
    )
  )
cat(sprintf("Finished recoding eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
```

## Merge behavioral and eye

```{r}
# merge the datasets based on trial number
cond_cs_cleaned <- eye_cond_cs %>%
  left_join(beh_cond, by = c('trial', 'cs', 'subject')) %>%
  mutate(time = time / 1000) %>% # ms to s
  select(subject, trial, time, pupil, cs, pos_item, incong_loc, iter, gazex, gazey, ia)

cat(sprintf("Merged cleaning eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
```

## CS Presentation

### Pupil Dilation

```{r}
# Window of interest (0.5-1.8s)
window_start = .5
window_end = 1.8
cat(sprintf("Looking at window of %s - %ss", window_start, window_end))
```

#### How does pupil changes after cs onset?

```{r}
# Calculate average pupil by time and cs
cond_cs_time <- cond_cs_cleaned %>%
  group_by(subject, time, cs) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE), .groups = 'drop')

# Visualize

# plot aesthetics
breaks_x <- seq(0, 2, by = 0.5)
labels_x <- ifelse(breaks_x %% 1 == 0, as.integer(breaks_x), as.character(breaks_x)) 
breaks_y <- seq(-1.5, 0.5, by = 0.5) # customize x, y-axis
colors <- c("alc" = "#e4b2db", "neut" = "#73cebc", "null" = "#0773a9")

p_cond_cs_time <- ggplot(cond_cs_time, aes(x = time, y = avg_pupil, color = cs)) +
    geom_line(size = 0.8) +
    labs(title = "Averaged Pupil Responses During CS Onset",
         x = "Time (s)",
         y = "Baseline-corrected pupil diamter (mm)") +
    scale_x_continuous(limits = c(0, 2), breaks = breaks_x, labels = labels_x) +
    scale_y_continuous(breaks = breaks_y) +
    scale_color_manual(values = colors) +
    geom_vline(xintercept = c(window_start, window_end)) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          panel.grid = element_blank(),
          axis.line = element_line(color="black")
    )
p_cond_cs_time
save_plot(file.path(output_path, "p_cond_cs_time.png"), p_cond_cs_time)
```

#### Comparisons in average pupil between CS

```{r}
# Calculate average pupil by cs and trial
avg_cond_cs <- cond_cs_cleaned %>%
  filter(between(time, window_start, window_end)) %>%
  group_by(trial, cs) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE))
  
p_avg_cond_cs <- ggplot(avg_cond_cs, aes(x = cs, y = avg_pupil, fill = cs)) +
  geom_violin(trim = FALSE) +
  scale_fill_manual(values = colors) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3, shape = 18) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, color = "black") +
  labs(title = "Distribution of Pupil Responses by CS",
       x = "CS type",
       y = "Average baseline-corrected pupil diameter (mm)") +
  theme_minimal() +
  theme(axis.line = element_line(color = "black"),
        panel.grid = element_blank())
p_avg_cond_cs
#save_plot(file.path(output_path, "p_avg_cond_cs.png"), p_avg_cond_cs)

```

#### How does pupil changes in perspective of learning quartiles?

```{r}
# Calculate the average pupil dilation for each quartile for each condition
avg_pupil_quart <- cond_cs_cleaned %>%
  group_by(subject, cs) %>%
  mutate(quartile = ntile(iter, 4)) %>%
  ungroup() %>%
  group_by(subject, cs, quartile, time) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE)) %>%
  ungroup()

# Visualize
p_cond_cs_quart <- ggplot(avg_pupil_quart, aes(x = time, y = avg_pupil, color = as.factor(cs), group = interaction(cs, quartile))) +
    geom_line(size = 0.8) +
    facet_wrap(~ quartile, labeller = label_value) +
    labs(title = "Pupil Dilation Over Time by Quartile",
         x = "Time after CS onset (s)",
         y = "Baseline-corrected pupil diameter (mm)",
         color = "cs") +
    scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
    scale_y_continuous(breaks = breaks_y) +
    scale_color_manual(values = colors) +
    geom_vline(xintercept = window_start) +
    geom_vline(xintercept = window_end) +
    theme_minimal() +
    theme(plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(color="black"))
p_cond_cs_quart
save_plot(file.path(output_path, "p_cond_cs_quart.png"), p_cond_cs_quart)
```

#### How does pupil changes as repetition unfolds?

```{r}
# divide into iterations
cond_cs_iter <- cond_cs_cleaned %>%
  group_by(subject, cs, iter) %>%
  filter(time >= window_start & time <= window_end) %>%
  summarize(avg_pupil = mean(pupil, na.rm = T))

# lm
model <- lm(avg_pupil ~ cs * iter, data = cond_cs_iter)
saveRDS(model, file = file.path(output_path, "lm_avgpupil_cs_iter.rds"))
  
anova_results <- anova(model)
print(anova_results)
  
p_lm <- plot_model(model, type = "eff", terms = c("iter", "cs"),
            title = "Average Pupil Diameter by Repetition",
            axis.title = c("Repetition", "Baseline-corrected pupil diameter (mm)")) +
  theme_minimal() +
  theme(axis.title = element_text(),
  axis.text = element_text(),
          plot.title = element_text(hjust = 0.5),
          legend.position = "bottom",
          legend.title = element_text(),
          legend.text = element_text(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(color = "black")) +
  scale_color_manual(values = colors)
p_lm
save_plot(file.path(output_path, "p_cond_us_rep.png"), p_lm)
```

#### LME

```{r}
# fit LME (linear mixed effects) models on pupil dilation
## How is average pupil diameter influenced by condition, quartile, and iteration?
analyze_linear_models <- function(data, output_path) {
  cat(sprintf("Analyzing linear models on pupil dilation for subject %s, task %s, phase %s...\n", subject, task, phase))
  model <- lm(avg_pupil ~ cs * iter, data = data)
  saveRDS(model, file = file.path(output_path, "lm_avgpupil_cs_iter.rds"))
  
  anova_results <- anova(model)
  print(anova_results)
  
  p_lm <- plot_model(model, type = "eff", terms = c("iter", "cs"),
            title = "Trend of Pupil Dilation by Repetition",
            axis.title = c("Repetition", "Baseline-corrected pupil diameter (mm)"))
  
  emm_results <- emmeans(model, pairwise ~ cs | iter, at = list(iter = c(1:40)))
  print(emm_results)
  
  emt_results <- test(emtrends(model, var = "iter", ~ cs))
  print(emt_results)
}
```

```{r}
# bar graph for average pupil dilation by quartile
p_cond_cs_avg_quart <- ggplot(pupil_windowed, aes(x = quartile, y = avg_pupil, fill = cs)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Average Pupil Diameter by Quartile (0.5s - 1.8s)",
        x = "Quartile",
        y = "Baseline-corrected pupil diameter (mm)",
        fill = "Condition") +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(color="black"))
save_plot(file.path(output_path, "p_cond_cs_avg_quart.png"), p_cond_cs_avg_quart)
```

## US Presentation

```{r}
# Load the preprocessed eye-tracking data
eye_cond_us_raw <- read_csv(input_path_eye, show_col_types = FALSE)

eye_cond_us <- eye_cond_us_raw %>% 
  select(subject = Subject,
         trial = TRIAL_INDEX,
         time = Time,
         pupil = Pupil,
         cs = cs,
         ia = RIGHT_INTEREST_AREA_ID,
         gazex = Gaze_X,
         gazey = Gaze_Y) %>%
  mutate(
    cs = case_when(
      cs %in% vcs_list[1] ~ vcs_ord[1],
      cs %in% vcs_list[2] ~ vcs_ord[2],
      cs %in% vcs_list[3] ~ vcs_ord[3],
      TRUE ~ cs
    )
  )
cat(sprintf("Finished recoding eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))

# merge the datasets based on trial number
cond_cleaned <- eye_cond_us %>%
  filter(ia == 1) %>%
  left_join(beh_cond, by = c('trial', 'cs', 'subject')) %>%
  mutate(time = time / 1000) %>% # ms to s
  select(subject, trial, time, pupil, cs, pos_item, incong_loc, iter, gazex, gazey, ia)

cat(sprintf("Merged cleaning eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
```

### Analysis and Plots

```{r}
# # US_ONSET: analyze baseline-corrected pupil dilation

cat(sprintf("Analyzing baseline_corrected pupil dilation for subject %s, task %s, phase %s...\n", subject, task, phase))

# within the window
window_start = .5
window_end = 1.8
cat(sprintf("Looking at the time window between %ss and %ss", window_start, window_end))

cond_us_time <- cond_cleaned %>%
  filter(incong_loc == "FALSE",
         cs != "null") %>%
  group_by(time, cs) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE, .groups = "drop"))

breaks_x <- seq(0, 2, by = 0.5)
labels_x <- ifelse(breaks_x %% 1 == 0, as.integer(breaks_x), as.character(breaks_x)) 
breaks_y <- seq(-1.5, 0.5, by = 0.5) # customize x, y-axis

p_cond_us_time <- ggplot(cond_us_time, aes(x = time, y = avg_pupil, color = cs)) +
  geom_line(size = 0.8) +
  labs(title = "Pupil Dilation Over Time in Congruent ROI",
       x = "Time after US onset (s)",
       y = "Baseline-corrected pupil diameter (mm)") +
  scale_x_continuous(limits = c(0, 2), breaks = breaks_x, labels = labels_x) +
  scale_y_continuous(breaks = breaks_y) +
  scale_color_manual(values = colors) + 
  geom_vline(xintercept = window_start) +
  geom_vline(xintercept = window_end) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"))
p_cond_us_time
save_plot(file.path(output_path, "p_cond_us_time.png"), p_cond_us_time)
  
# avg pupil for each us
avg_cond_us <- cond_cleaned %>%
  filter(between(time, window_start, window_end),
         cs != "null") %>%
  group_by(trial, cs) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE),
              .groups = "drop")

p_cond_us_avg <- ggplot(avg_cond_us, aes(x = cs, y = avg_pupil, fill = cs)) +
  geom_violin(trim = FALSE) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3, shape = 18) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2, color = "black") +
  labs(title = "Distribution of Pupil Responses Following US Onset",
       x = "CS type",
       y = "Baseline-corrected pupil diameter (mm)") +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(plot.title.position = "plot",
          plot.title = element_text(hjust = 0.2),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(color="black"))
p_cond_us_avg
save_plot(file.path(output_path, "p_cond_us_avg.png"), p_cond_us_avg)

cond_us_iter <- cond_cleaned %>%
  filter(incong_loc == "FALSE",
         cs != "null") %>%
  group_by(time, cs, iter) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE, .groups = "drop"))

model <- lm(avg_pupil ~ cs * iter, data = cond_us_iter)
p_lm <- plot_model(model, type = "eff", terms = c("iter", "cs"),
            title = "Trend of Pupil Dilation by Repetition",
            axis.title = c("Repetition", "Baseline-corrected pupil diameter (mm)")) +
  theme_minimal() +
  theme(axis.title = element_text(),
  axis.text = element_text(),
          plot.title = element_text(hjust = 0.5),
          legend.position = "bottom",
          legend.title = element_text(),
          legend.text = element_text(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(color = "black")) +
  scale_color_manual(values = colors)
save_plot(file.path(output_path, "p_cond_us_rep.png"), p_lm)
p_lm
```

# Generalization Task

## Load behavioral data and eye-traking data

```{r}
subject <- c("L003")  # Replace with actual subject
task <- list(c("generalization")) # Replace with actual task (conditioning/generalization/csmem/usmem)
task_abbr <- list(c("gen")) # abbreviation of the task for naming the variable (cond/gen/csmem/usmem)
phase <- list(c("gs_onset")) # Replace with actual phase (cs_onset/anticipation/us_onset/gs_onset)
load_beh(subject, task, phase, task_abbr)

eye_raw <- read_csv(input_path_eye, show_col_types = FALSE)
assign(sprintf("eye_%s_raw", task_abbr), eye_raw, envir = .GlobalEnv)
cat(sprintf("Finished loading eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
```

## Recode behavior

```{r}
beh_gen <- beh_gen_raw %>%
  filter(block != "prac")

# Figure out which image is which GS type
gen_vcs_list <- c(unique(beh_gen[grepl('null', beh_gen$category),]$gs),
              unique(beh_gen[grepl('novel', beh_gen$category),]$gs),
              unique(beh_gen[grepl('neut', beh_gen$category),]$gs),
              unique(beh_gen[grepl('alc', beh_gen$category),]$gs)) # should be 25 of them
gen_vcs_ord <- c('null', 'novel', 'neut', 'alc')

beh_gen <- beh_gen %>%
  mutate(
    cs = case_when(
      gs %in% gen_vcs_list[c(1, 2, 3, 4, 5, 6, 7)] ~ gen_vcs_ord[1],
      gs %in% gen_vcs_list[c(8, 9, 10, 11)] ~ gen_vcs_ord[2],
      gs %in% gen_vcs_list[c(12, 13, 14, 15, 16, 17, 18)] ~ gen_vcs_ord[3],
      gs %in% gen_vcs_list[c(19, 20, 21, 22, 23, 24, 25)] ~ gen_vcs_ord[4],
      TRUE ~ cs)
  ) %>%
  select(subject, phase, trial, category, gs_deg, probe_pos, incong_loc, gs, cs)
```

## Recode eye

```{r}
eye_gen <- eye_gen_raw %>%
  mutate(
    cs = case_when(
      gs %in% gen_vcs_list[c(1, 2, 3, 4, 5, 6, 7)] ~ gen_vcs_ord[1],
      gs %in% gen_vcs_list[c(8, 9, 10, 11)] ~ gen_vcs_ord[2],
      gs %in% gen_vcs_list[c(12, 13, 14, 15, 16, 17, 18)] ~ gen_vcs_ord[3],
      gs %in% gen_vcs_list[c(19, 20, 21, 22, 23, 24, 25)] ~ gen_vcs_ord[4],
      TRUE ~ cs),
  ) %>%
  select(subject = Subject,
         trial = TRIAL_INDEX,
         time = Time,
         pupil = Pupil,
         cs = cs,
         gs =gs,
         diss = gs_deg
         ) %>%
  mutate(time = time / 1000)
cat(sprintf("Finished cleaning eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
```

#### Pupil dilation in function of GS

```{r}
gen_pupil_full <- eye_gen %>%
  group_by(time, cs) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE),
            .groups = "drop")

p_gen_pupil_full <- ggplot(gen_pupil_full, aes(x = time, y = avg_pupil, color = cs)) +
  geom_line(size = 0.8) +
  labs(title = "Pupil Dilation Over Time",
       x = "Time (s)",
       y = "Baseline-corrected pupil diameter (mm)") +
  scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
  scale_y_continuous(breaks = breaks_y) +
  scale_color_manual(values = colors) +
  geom_vline(xintercept = window_start) +
  geom_vline(xintercept = window_end) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color="black")
  )

# Print the plot
p_gen_pupil_full
save_plot(file.path(output_path, "p_gen_pupil_full.png"), p_gen_pupil_full)


# average
gen_avg_win <- gen_pupil_full %>%
    filter(between(time, window_start, window_end)) %>% 
    group_by(cs) %>%
    summarize(avg_pupil = mean(avg_pupil, na.rm = TRUE),
              .groups = "drop")

p_gen_avg_win <- ggplot(gen_avg_win, aes(x = cs, y = avg_pupil, fill = cs)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7) +
    labs(title = "Average Pupil Diameter (0.5s - 1.8s)",
         x = "Condition",
         y = "Average baseline-corrected pupil diameter (mm)",
         fill = "Condition") +
    scale_fill_manual(values = colors) +
    theme_minimal()
p_gen_avg_win
save_plot(file.path(output_path, "p_gen_avg_win.png"), p_gen_avg_win)
```

#### Pupil dilation in function of GS and perceptual dissimilarity

```{r}
gen_pupil_diss <- eye_gen %>%
  group_by(time, cs, diss) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE))

# Convert 'diss' and 'cs' to factor
gen_pupil_diss$diss <- as.factor(gen_pupil_diss$diss)
# Rename to orignal and novel cs
levels(gen_pupil_diss$diss)[levels(gen_pupil_diss$diss) == "0"] <- "Original"
levels(gen_pupil_diss$diss)[levels(gen_pupil_diss$diss) == "99"] <- "Novel"
gen_pupil_diss$cs <- as.factor(gen_pupil_diss$cs)

# Create the plot
p_gen_pupil_dis <- ggplot(gen_pupil_diss, aes(x = time, y = avg_pupil, color = cs)) +
  geom_line(size = 0.8) +
  facet_wrap(~ diss, scales = "free_x", ncol = 4) +
  labs(title = "Pupil Dilation Over Time by Dissimilarity",
       x = "Time (s)",
       y = "Baseline-corrected pupil diameter (mm)") +
  scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
  scale_y_continuous(breaks = breaks_y) +
  scale_color_manual(values = colors) +
  geom_vline(xintercept = window_start) +
  geom_vline(xintercept = window_end) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color="black")
  )

# Print the plot
p_gen_pupil_dis
save_plot(file.path(output_path, "p_gen_pupil_dis.png"), p_gen_pupil_dis, width = 18, height = 12)

# visualize avg pupil per dissimilarity
# note: will want to group by subject, x-axis and color terms, then compute avg
p_gen_pupilavg_dis <- ggplot(gen_pupil_diss %>%
                            filter(time >= window_start & time <= window_end), 
                          aes(x = diss, y = avg_pupil, color = cs, group = cs)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Pupil Dilation by Dissimilarity",
       x = "Dissimilarity",
       y = "Average baseline-corrected pupil diameter (mm)") +
  scale_color_manual(values = colors) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color="black"))

save_plot(file.path(output_path, "p_gen_pupilavg_dis.png"), p_gen_pupilavg_dis)
# with model fit
p_gen_pupilavg_dis_fit <- ggplot(gen_pupil_diss %>%
                            filter(time >= window_start & time <= window_end), 
                          aes(x = diss, y = avg_pupil, color = cs, group = cs, fill = cs)) +
  geom_smooth(method = "loess") +
  labs(title = "Pupil Dilation by Dissimilarity",
       x = "Dissimilarity (degree)",
       y = "Average baseline-corrected pupil diameter (mm)") +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color="black"))
save_plot(file.path(output_path, "p_gen_pupilavg_dis_fit.png"), p_gen_pupilavg_dis_fit)
```

## Gaze data

```{r}
gaze_x <- eye_cond_raw %>%
  select(subject = Subject,
         trial = TRIAL_INDEX,
         time = Time,
         pupil = Pupil,
         cs = cs,
         gazex = Gaze_X,
         ia = RIGHT_INTEREST_AREA_ID) %>%
  mutate(
    cs = case_when(
      cs %in% vcs_list[1] ~ vcs_ord[1],
      cs %in% vcs_list[2] ~ vcs_ord[2],
      cs %in% vcs_list[3] ~ vcs_ord[3],
      TRUE ~ cs
    )
  )
gaze_x <- gaze_x %>%
  group_by(time, cs) %>%
    summarize(avg_gazex = mean(gazex, na.rm = TRUE))

p_cond_gaze <- ggplot(gaze_x, aes(x = time, y = avg_gazex, color = cs)) +
  geom_line(size = 0.8) +
  labs(title = "Gaze Direction Over Time",
       x = "Time (s)",
       y = "Gaze x") +
  scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
  scale_y_continuous(breaks = breaks_y) +
  scale_color_manual(values = colors) +
  geom_vline(xintercept = window_start) +
  geom_vline(xintercept = window_end) +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color="black")
  )
p_cond_gaze

```

# Backup code

```{r}
# gen_pupil_diss$diss <- as.factor(gen_pupil_diss$diss)
# 
# # Define colors for each level of 'diss'
# diss_colors <- c("Original" = "red", "10" = "blue", "20" = "green", 
#                  "30" = "purple", "40" = "orange", "50" = "pink",
#                  "60" = "brown", "Novel" = "grey")
# 
# # Create the plot
# p_gen_pupil_all <- ggplot(gen_pupil_diss, aes(x = time, y = avg_pupil, color = diss)) +
#   geom_line(size = 0.8) +
#   facet_wrap(~ cs, scales = "free_x", ncol = 4) +
#   labs(title = "Overall Pupil Dilation Over Time",
#         x = "Time (s)",
#         y = "Baseline-corrected pupil diameter (mm)") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = diss_colors) +
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black")
#   )
# 
# # Print the plot
# p_gen_pupil_all


# # overall pupil by congruency
# pupil_full_cong <- cond_cleaned %>%
#   filter(cs != "null") %>%
#   group_by(time, cs, pos_item, incong_loc, iter) %>%
#   summarize(avg_pupil = mean(pupil, na.rm = TRUE), .groups = "drop") %>%
#   mutate(cong = if_else(incong_loc == "FALSE", "congruent", "incongruent"))
# 
# # find null separately
# null_left <- read_csv("/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab/eye/preprocessed/L003_null_left.csv")
# null_right <- read_csv("/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab/eye/preprocessed/L003_null_right.csv")
# 
# null_left <- null_left %>%
#   mutate(ia = "left")
# null_right <- null_right %>%
#   mutate(ia = "right")
# 
# null <- bind_rows(null_left, null_right)
# null <- null %>%
#   select(subject = Subject,
#          trial = TRIAL_INDEX,
#          time = Time,
#          pupil = Pupil,
#          cs = cs,
#          ia = ia) %>%
#   mutate(
#     cstype = case_when(
#       cs == "VCS_10.jpg" ~ "null"),
#     cs = cstype,
#     time = time / 1000)
# 
# pupil_null_cong <- null %>%
#   mutate(cong = if_else(ia == "left", "congruent", "incongruent")) %>%
#   group_by(time, cong, cs, trial) %>%
#   summarize(avg_pupil = mean(pupil, na.rm = TRUE), .groups = "drop")
# 
# # merge all CS types
# us_time_cong <- bind_rows(pupil_full_cong, pupil_null_cong)
# 
# 
# # Create interaction term for the legend
# us_time_cong$Congruency <- interaction(us_time_cong$cs, us_time_cong$cong)

# Plot with separate graphs for each congruency
# p_pupil_full_cong <- ggplot(us_time_cong, aes(x = time, y = avg_pupil, color = cs)) + # Change here
#   geom_line(size = 0.8) +
#   facet_wrap(~ cong, scales = "free_y") + # Facet by congruency
#   labs(title = "Pupil Dilation Over Time in Congruent ROI",
#        x = "Time after US onset (s)",
#        y = "Baseline-corrected pupil diameter (mm)") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = colors) + 
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color = "black"))
# save_plot(file.path(output_path, "p_pupil_full_cong.png"), p_pupil_full_cong)

# # window average by congruency
# pupil_avg_cong <- us_time_cong %>%
#   filter(between(time, window_start, window_end)) %>%
#   group_by(cs, cong) %>%
#   summarize(avg_pupil = mean(avg_pupil, na.rm = TRUE)) %>%
#   ungroup()
# pupil_avg_cong
# 
# # Define colors
# intercolors <- c("alc.congruent" = "#e4b2db", "neut.congruent" = "#73cebc", "null.congruent" = "#0773a9",
#             "alc.incongruent" = "#cc99c2", "neut.incongruent" = "#66c2a4", "null.incongruent" = "#055a86")
# pupil_avg_cong$cong <- factor(pupil_avg_cong$cong, levels = c("congruent", "incongruent"))
# pupil_avg_cong$Congruency <- interaction(pupil_avg_cong$cs, pupil_avg_cong$cong)
# legend_lab <- c("alc.congruent" = "congruent",
#                 "alc.incongruent" = "incongruent",
#                 "neut.congruent" = "congruent",
#                 "neut.incongruent" ="incongruent",
#                 "null.congruent" = "congruent",
#                 "null.incongruent" = "incongruent")
# 
# # Convert "cong" to a factor and create an interaction term for coloring
# pupil_avg_cong$cong <- factor(pupil_avg_cong$cong, levels = c("congruent", "incongruent"))
# pupil_avg_cong$Congruency <- interaction(pupil_avg_cong$cs, pupil_avg_cong$cong)
# 
# p_avg_pupil_cong <- ggplot(pupil_avg_cong, aes(x = cs, y = avg_pupil, fill = Congruency)) +
#   geom_bar(stat = "identity", position = "dodge", width = 0.7) +
#   labs(title = "Average Pupil Diameter by Congruency (0.5s - 1.8s)",
#        x = "Condition",
#        y = "Average baseline-corrected pupil diameter (mm)") +
#   scale_fill_manual(values = intercolors, labels = legend_lab) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black"))
# save_plot(file.path(output_path, "p_avg_pupil_cong.png"), p_avg_pupil_cong)
# ```
# 
# #### Quartile \~ Congruency
# 
# ```{r}
#   # recode
#   pupil_null_cong <- pupil_null_cong %>%
#     mutate(quartile = ntile(iter, 4)) %>%
#     ungroup()
# 
#   pupil_csneut_cong <- pupil_full_cong %>%
#     mutate(quartile = ntile(iter, 4)) %>%
#     ungroup() %>%
#     select(time, cong, cs, avg_pupil, iter, quartile)
#   
#   pupil_null_cong$trial <- NULL
#   allcs_cong <- bind_rows(pupil_null_cong, pupil_csneut_cong)
#   
#   # calculate the average pupil dilation for each quartile for each condition
#   
#   avg_cong_quart <- allcs_cong %>%
#     filter(between(time, window_start, window_end)) %>%
#     group_by(cs, quartile, cong) %>%
#     summarize(avg_pupil = mean(avg_pupil, na.rm = TRUE),
#               .groups = "drop")
#     
#   # bar graph for average pupil dilation between 0.5~1.8s
#   p_avg_cong_quart <- ggplot(avg_cong_quart, aes(x = quartile, y = avg_pupil, fill = cs)) +
#     geom_bar(stat = "identity", position = "dodge", width = 0.7) +
#     facet_wrap(~cong, scales = "free_y") +
#     labs(title = "Average Pupil Diameter by Quartile and Congruency",
#          x = "Quartile",
#          y = "Average baseline-corrected pupil diameter (mm)",
#          fill = "Condition") +
#     scale_fill_manual(values = colors) +
#     theme_minimal() +
#     theme(plot.title.position = "plot",
#           plot.title = element_text(hjust = 0.5),
#           panel.grid.major = element_blank(),
#           panel.grid.minor = element_blank(),
#           axis.line = element_line(color="black"))
# 
#   print(p_avg_cong_quart)
#   save_plot(file.path(output_path, "p_avg_cong_quart.png"), p_avg_cong_quart)
#   
# 
#   
# # divide into iterations
#   pupil_iter <- allcs_cong %>%
#     group_by(cs, iter, cong) %>%
#     filter(time >= window_start & time <= window_end) %>%
#     summarize(avg_pupil = mean(avg_pupil, na.rm = T))
# 
#   avg_cong_quart$quartile <- factor(avg_cong_quart$quartile,
#                                      levels = 1:4,
#                                      labels = c("Q1: Repetition 1~10", "Q2: Repetition 11~20", "Q3: Repetition 21~30", "Q4: Repetition 31~40"))
# 

```
