---
title: "AlcGen Pupil Data Analysis"
author: "Zihan Bai"
date: "`r Sys.Date()`"
description: "This script analyze pupil data for the AlcGen study. It reads preprocesed data, performs cleaning and analysis, and outputs figures"
output: html_document
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages("pacman)
library(pacman)
p_load(readr, dplyr, ggplot2, PupillometryR, PupilPre, shiny, tidyverse, stringr, emmeans, sjPlot, nlme, patchwork, purrr, ggpubr, lme4)
```

# Conditioning Task

# Initial function to load behavioral data - call it later

```{r}
# Function to process one single subject
process_subject <- function(subject, task, phase, base_path, output_path) {
  
    subID <- subject[1]
    drink_mapping <- c('L' = 'light', 'H' = 'heavy')
    # Load behavioral data
    input_path_beh <- file.path(base_path, subject)
    beh_pattern <- sprintf("%s_task_%s_events_.*\\.csv$", subject, task)
    beh_files <- list.files(input_path_beh, pattern = beh_pattern, full.names = TRUE)
    
    if (length(beh_files) != 1) {
        stop("Either no or multiple behavioral data files found. Please check the directory and pattern.")
    }

    beh_cond <- read_csv(beh_files[1], show_col_types = FALSE)
    beh_cond <- beh_cond %>%
      mutate(subject = substr(subject, 1, 4))
    
    # Recode behavior
    beh_cond <- beh_cond %>%
      mutate(
        stim = case_when(
          category == "neut" ~ "neut",
          category == "alc" ~ "alc",
          category == "null" ~ "null",
          category == "null_alc_surprise" ~ "null",
          category == "null_neut_surprise" ~ "null",
          category == "alc_null_surprise" ~ "alc",
          category == "neut_null_surprise" ~ "neut"),
        incong_loc = case_when(
          category %in% c("alc_null_surprise", "null_alc_surprise", 
                          "null_neut_surprise", "neut_null_surprise") ~ TRUE, # ignore 30% trials while congruent locs contain nothing
          TRUE ~ incong_loc)
      )

    vcs_list <- c(unique(beh_cond[grepl('neut', beh_cond$stim),]$cs),
                  unique(beh_cond[grepl('alc', beh_cond$stim),]$cs),
                  unique(beh_cond[grepl('null', beh_cond$stim),]$cs))
    vcs_ord <- c('neut', 'alc', 'null')

    # Find the number of exposure ppl have seen for each CS type
    beh_cond <- beh_cond %>%
      mutate(
        cs = case_when(
          cs %in% vcs_list[1] ~ vcs_ord[1],
          cs %in% vcs_list[2] ~ vcs_ord[2],
          cs %in% vcs_list[3] ~ vcs_ord[3],
          TRUE ~ cs)) %>%
      arrange(trial) %>%
      group_by(cs) %>%
      mutate(iter = row_number()) %>%
      ungroup()
    
    first_char <- substr(subID, 1, 1)  # Extract the first character of subID
    beh_cond$drink_2grp <- drink_mapping[first_char]
    
    # Load eye-tracking data
    input_path_eye <- file.path(base_path, "eye", "preprocessed", subject, sprintf("%s_%s_%s_preprocessed.csv", subject, task, phase))
    eye_cond_cs <- read_csv(input_path_eye, show_col_types = FALSE)

    # Recode eye-tracking data
    eye_cond_cs <- eye_cond_cs %>%
      mutate(cs = case_when(
               cs %in% vcs_list[1] ~ vcs_ord[1],
               cs %in% vcs_list[2] ~ vcs_ord[2],
               cs %in% vcs_list[3] ~ vcs_ord[3],
               TRUE ~ cs)) %>%
      mutate(subject = substr(subject, 1, 4)) %>%
      select(subject = subject,
             trial = TRIAL_INDEX,
             time = Time,
             pupil = Pupil,
             cs = cs,
             ia = RIGHT_INTEREST_AREA_ID,
             gazex = Gaze_X,
             gazey = Gaze_Y)
      
    
    # Merge behavioral and eye data
    merged_subject <- eye_cond_cs %>%
      left_join(beh_cond, by = c('trial', 'cs', 'subject')) %>%
      mutate(time = time / 1000) %>%
      select(subject, trial, time, pupil, cs, pos_item, incong_loc, iter, gazex, gazey, ia, drink_2grp) %>%
      filter(between(time, 0, 2))
    
    return(merged_subject)
}

# # What is the congruent ROI for each CS type for this subject?
# cong_ROI <- beh_cond %>%
#   select(cs, incong_loc, pos_item) %>%
#   filter(cs != "null" & incong_loc == "FALSE") %>%
#   group_by(cs) %>%
#   summarize(cong_ROI = unique(pos_item)) %>%
#   ungroup() %>%
#   print()

```

## Recode eye

## Merge behavioral and eye

```{r}
# Define subjects, task, phase, and paths
subject <- c('L003', 'L004', 'L006', 'L009')
# 
task <- "conditioning"
phase <- "cs_onset"
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"
output_path <- file.path(base_path, "eye", "results")

# Use lapply to process each subject
all_subjects_data <- lapply(subject, process_subject, task, phase, base_path, output_path)

# Combine data from all subjects into a single dataframe
combined_cond <- bind_rows(all_subjects_data)
write_csv(combined_cond, file.path(output_path, "combined_cond_cs.csv"))

```

## CS Presentation

### Pupil Dilation

```{r}
# Window of interest (0.5-1.8s)
window_start = 0.5
window_end = 1.8
cat(sprintf("Looking at window of %s - %ss", window_start, window_end))

# plot aesthetics
breaks_x <- seq(0, 2, by = 0.5)
labels_x <- ifelse(breaks_x %% 1 == 0, as.integer(breaks_x), as.character(breaks_x)) 
breaks_y <- seq(-1.5, 0.5, by = 0.25) # customize x, y-axis
colors <- c("alc" = "#e4b2db", "neut" = "#73cebc", "null" = "#0773a9")
```
#### How does pupil changes after cs onset?

```{r}
# calculation
cond_cs_time <- combined_cond %>%
  filter(between(time, window_start, window_end)) %>%
  group_by(drink_2grp, cs, time) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE),
            n = n(),
            sd = sd(pupil, na.rm = TRUE),
            .groups = 'drop') %>%
  mutate(se = sd / sqrt(n))

# sample size?
num_subjects <- combined_cond %>% group_by(drink_2grp) %>% summarize(n = n_distinct(subject), .groups = 'keep')
num_subjects$facet_label <- paste(num_subjects$drink_2grp, "(n=", num_subjects$n, ")", sep="")
cond_cs_time <- cond_cs_time %>% left_join(num_subjects, by = "drink_2grp")

# plot
p_cond_cs_time <- ggplot(cond_cs_time, aes(x = time, y = avg_pupil, group = interaction(cs, drink_2grp))) +
  geom_ribbon(aes(fill = cs, ymin = avg_pupil - se, ymax = avg_pupil + se), alpha = 0.2) +
  geom_line(aes(color = cs), size = 1) +
  facet_wrap(~facet_label) +
  labs(title = "Averaged Pupil Responses During CS Onset", x = "Time after cs onset (s)", y = "Baseline-corrected pupil diameter (mm)") +
  scale_x_continuous(limits = c(0,2), breaks = breaks_x, labels = labels_x) +
  scale_y_continuous(breaks = breaks_y) +
  scale_color_manual(values = colors) + scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black"), legend.position = "bottom") +
  geom_vline(xintercept = window_start, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = window_end, linetype = "dashed", color = "grey")


print(p_cond_cs_time)
ggsave(file.path(output_path, "p_cond_cs_time.pdf"), p_cond_cs_time, width = 6, height = 8)
```

#### Comparisons in average pupil between CS

```{r}
p_cond_cs_avg <- ggplot(cond_cs_time, aes(x = interaction(cs, drink_2grp), y = avg_pupil, fill = cs)) +
  geom_boxplot(alpha = 0.5) +
  geom_violin(trim = FALSE, alpha = 0.3) +
  labs(title = "Distribution of Averaged Pupil Responses",
       x = "CS and Drinking Group",
       y = "Baseline-corrected pupil diameter (mm)") +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black"), legend.position = "bottom")

# Print the plot
print(p_cond_cs_avg)

ggsave(file.path(output_path, "p_cond_cs_avg.pdf"), p_cond_cs_avg)

# model
model <- lm(avg_pupil ~ cs * drink_2grp, data = cond_cs_time)
summary(model)

# Pupil dilation varies significantly between different drinking groups.
# The type of conditioned stimulus (CS category) alone does not significantly change pupil dilation when controlling for drinking group.
# The interaction between the type of CS and the drinking group does not significantly influence pupil dilation.

model_aov <- aov(avg_pupil ~ cs * drink_2grp, data = cond_cs_time)
summary(model_aov)
```
#### Does pupil dilation vary with trial repetition?
```{r}
# divide into iterations
cond_cs_iter <- combined_cond %>%
  filter(between(time, window_start, window_end)) %>%
  group_by(drink_2grp, cs, iter) %>%
  summarize(avg_pupil = mean(pupil, na.rm = T))

# lm
lm_cond_cs <- lme(avg_pupil ~ cs * iter * drink_2grp + random=~1|subject, data = cond_cs_iter)
anova_results <- anova(lm_cond_cs)
print(anova_results)
capture.output(summary(anova_results), file = "anova_results.txt")
  
p_lm_cond_cs <- plot_model(lm_cond_cs, type = "eff", terms = c("iter", "cs", "drink_2grp"),
            title = "Average Pupil Reponses by Repetition",
            axis.title = c("Repetition", "Baseline-corrected pupil diameter (mm)")) +
  theme_minimal() + theme(axis.title = element_text(), axis.text = element_text(),plot.title = element_text(hjust = 0.5), legend.position = "bottom",legend.title = element_text(), legend.text = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black")) +
  scale_color_manual(values = colors)
p_lm_cond_cs
ggsave(file.path(output_path, "p_lm_cond_cs.pdf"), p_lm_cond_cs, width = 8, height = 6)
```


#### Granularity: Does pupil dilation vary with trial quartiles?

```{r}
# Calculate the average pupil dilation for each quartile for each condition
cond_cs_quar <- combined_cond %>%
  group_by(drink_2grp, cs, iter) %>%
  mutate(quartile = ntile(iter, 4)) %>%
  ungroup() %>%
  group_by(drink_2grp, cs, quartile) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE))

p_cond_cs_quar <- ggplot(cond_cs_quar, aes(x = quartile, y = avg_pupil, color = cs)) +
  geom_point() +
  facet_wrap(~drink_2grp) +
  geom_smooth(method = "loess", formula = y ~ x, se = FALSE) +
  labs(
    title = "Average Pupil Dilation by Quartile",
    x = "Quartile",
    y = "Baseline-corrected pupil diameter (mm)"
  ) +
  scale_color_manual(values = colors) +
  theme_minimal() + theme(axis.title = element_text(), axis.text = element_text(),plot.title = element_text(hjust = 0.5), legend.position = "bottom",legend.title = element_text(), legend.text = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black"))

ggsave(file.path(output_path, "p_cond_cs_quar.pdf"), p_cond_cs_quar, width = 8, height = 6)
```



<!-- #### LME -->

<!-- ```{r} -->
<!-- # fit LME (linear mixed effects) models on pupil dilation -->
<!-- ## How is average pupil diameter influenced by condition, quartile, and iteration? -->

<!-- model <- lm(avg_pupil ~ cs * iter, data = data) -->
<!-- saveRDS(model, file = file.path(output_path, "lm_avgpupil_cs_iter.rds")) -->

<!-- anova_results <- anova(model) -->
<!-- print(anova_results) -->

<!-- p_lm <- plot_model(model, type = "eff", terms = c("iter", "cs"), -->
<!--             title = "Trend of Pupil Dilation by Repetition", -->
<!--             axis.title = c("Repetition", "Baseline-corrected pupil diameter (mm)")) -->

<!-- emm_results <- emmeans(model, pairwise ~ cs | iter, at = list(iter = c(1:40))) -->
<!-- print(emm_results) -->

<!-- emt_results <- test(emtrends(model, var = "iter", ~ cs)) -->
<!-- print(emt_results) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # bar graph for average pupil dilation by quartile -->
<!-- p_cond_cs_avg_quart <- ggplot(pupil_windowed, aes(x = quartile, y = avg_pupil, fill = cs)) + -->
<!--   geom_bar(stat = "identity", position = "dodge", width = 0.7) + -->
<!--   labs(title = "Average Pupil Diameter by Quartile (0.5s - 1.8s)", -->
<!--         x = "Quartile", -->
<!--         y = "Baseline-corrected pupil diameter (mm)", -->
<!--         fill = "Condition") + -->
<!--   scale_fill_manual(values = colors) + -->
<!--   theme_minimal() + -->
<!--   theme(plot.title.position = "plot", -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           panel.grid.major = element_blank(), -->
<!--           panel.grid.minor = element_blank(), -->
<!--           axis.line = element_line(color="black")) -->
<!-- save_plot(file.path(output_path, "p_cond_cs_avg_quart.png"), p_cond_cs_avg_quart) -->
<!-- ``` -->

## US Presentation

```{r}
# Define subjects, task, phase, and paths
subject <- c('L003', 'L004', 'L006', 'L009', 'H001')
task <- "conditioning"
phase <- "us_onset"
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"
output_path <- file.path(base_path, "eye", "results")

# Use lapply to process each subject
all_subjects_data <- lapply(subject, process_subject, task, phase, base_path, output_path)
all_subjects_data <- lapply(all_subjects_data, function(df) {
  df$ia <- as.character(df$ia)
  return(df)
})

# Combine data from all subjects into a single dataframe
combined_cond_us <- bind_rows(all_subjects_data)
write_csv(combined_cond_us, file.path(output_path, "combined_cond_us.csv"))
```

#### Increased pupil dilation assocaited with US after conditioning?
```{r}
cond_us_time <- combined_cond_us %>%
  filter(between(time, window_start, window_end),
         incong_loc == "FALSE",
         cs != "null") %>%
  group_by(drink_2grp, cs, time) %>%
  summarize(avg_pupil = mean(pupil, na.rm = TRUE),
            n = n(),
            sd = sd(pupil, na.rm = TRUE),
            .groups = 'drop') %>%
  mutate(se = sd / sqrt(n))

# sample size?
num_subjects <- combined_cond_us %>% group_by(drink_2grp) %>% summarize(n = n_distinct(subject), .groups = 'keep')
num_subjects$facet_label <- paste(num_subjects$drink_2grp, "(n=", num_subjects$n, ")", sep="")
cond_us_time <- cond_us_time %>% left_join(num_subjects, by = "drink_2grp")

# plot
p_cond_us_time <- ggplot(cond_us_time, aes(x = time, y = avg_pupil, group = interaction(cs, drink_2grp))) +
  geom_ribbon(aes(fill = cs, ymin = avg_pupil - se, ymax = avg_pupil + se), alpha = 0.2) +
  geom_line(aes(color = cs), size = 1) +
  facet_wrap(~facet_label) +
  labs(title = "Averaged Pupil Responses During US Onset", subtitle = "Congruent Location", x = "Time after us onset (s)", y = "Baseline-corrected pupil diameter (mm)") +
  scale_x_continuous(limits = c(0,2), breaks = breaks_x, labels = labels_x) + scale_y_continuous(breaks = breaks_y) +
  scale_color_manual(values = colors) + scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black"), legend.position = "bottom") +
  geom_vline(xintercept = window_start, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = window_end, linetype = "dashed", color = "grey")

print(p_cond_us_time)

ggsave(file.path(output_path, "p_cond_us_time.pdf"), p_cond_us_time, width = 6, height = 8)
```

#### Comparisons in average pupil between US
```{r}
p_cond_us_avg <- ggplot(cond_us_time, aes(x = interaction(cs, drink_2grp), y = avg_pupil, fill = cs)) +
  geom_boxplot(alpha = 0.5) +
  geom_violin(trim = FALSE, alpha = 0.3) +
  labs(title = "Distribution of Averaged Pupil Responses",
       x = "CS and Drinking Group",
       y = "Baseline-corrected pupil diameter (mm)") +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black"), legend.position = "bottom")

# Print the plot
print(p_cond_us_avg)
ggsave(file.path(output_path, "p_cond_us_avg.pdf"), p_cond_us_avg, width = 6, height = 6)
```


```{r}
cond_us_iter <- combined_cond_us %>%
  filter(between(time, window_start, window_end),
         incong_loc == "FALSE",
         cs != "null") %>%
  group_by(drink_2grp, cs, iter) %>%
  summarize(avg_pupil = mean(pupil, na.rm = T))

lm_cond_us <- lm(avg_pupil ~ cs * iter * drink_2grp, data = cond_us_iter)
anova_cond_us <- anova(lm_cond_us)
print(anova_cond_us)

p_lm_cond_us <- plot_model(lm_cond_us, type = "eff", terms = c("iter", "cs", "drink_2grp"), title = "Average Pupil Reponses by Repetition",axis.title = c("Repetition", "Baseline-corrected pupil diameter (mm)")) +
  theme_minimal() + theme(axis.title = element_text(), axis.text = element_text(),plot.title = element_text(hjust = 0.5), legend.position = "bottom",legend.title = element_text(), legend.text = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black")) +
  scale_color_manual(values = colors)

p_lm_cond_us
ggsave(file.path(output_path, "p_lm_cond_us.pdf"), p_lm_cond_us, width = 6, height = 6)
```




# Generalization Task
```{r}
# 
# ## Load behavioral data and eye-traking data
# 
# ```{r}
# subject <- c("L003")  # Replace with actual subject
# task <- list(c("generalization")) # Replace with actual task (conditioning/generalization/csmem/usmem)
# task_abbr <- list(c("gen")) # abbreviation of the task for naming the variable (cond/gen/csmem/usmem)
# phase <- list(c("gs_onset")) # Replace with actual phase (cs_onset/anticipation/us_onset/gs_onset)
# load_beh(subject, task, phase, task_abbr)
#
# eye_raw <- read_csv(input_path_eye, show_col_types = FALSE)
# assign(sprintf("eye_%s_raw", task_abbr), eye_raw, envir = .GlobalEnv)
# cat(sprintf("Finished loading eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
# ```
#
# ## Recode behavior
#
# ```{r}
# beh_gen <- beh_gen_raw %>%
#   filter(block != "prac")
#
# # Figure out which image is which GS type
# gen_vcs_list <- c(unique(beh_gen[grepl('null', beh_gen$category),]$gs),
#               unique(beh_gen[grepl('novel', beh_gen$category),]$gs),
#               unique(beh_gen[grepl('neut', beh_gen$category),]$gs),
#               unique(beh_gen[grepl('alc', beh_gen$category),]$gs)) # should be 25 of them
# gen_vcs_ord <- c('null', 'novel', 'neut', 'alc')
#
# beh_gen <- beh_gen %>%
#   mutate(
#     cs = case_when(
#       gs %in% gen_vcs_list[c(1, 2, 3, 4, 5, 6, 7)] ~ gen_vcs_ord[1],
#       gs %in% gen_vcs_list[c(8, 9, 10, 11)] ~ gen_vcs_ord[2],
#       gs %in% gen_vcs_list[c(12, 13, 14, 15, 16, 17, 18)] ~ gen_vcs_ord[3],
#       gs %in% gen_vcs_list[c(19, 20, 21, 22, 23, 24, 25)] ~ gen_vcs_ord[4],
#       TRUE ~ cs)
#   ) %>%
#   select(subject, phase, trial, category, gs_deg, probe_pos, incong_loc, gs, cs)
# ```
#
# ## Recode eye
#
# ```{r}
# eye_gen <- eye_gen_raw %>%
#   mutate(
#     cs = case_when(
#       gs %in% gen_vcs_list[c(1, 2, 3, 4, 5, 6, 7)] ~ gen_vcs_ord[1],
#       gs %in% gen_vcs_list[c(8, 9, 10, 11)] ~ gen_vcs_ord[2],
#       gs %in% gen_vcs_list[c(12, 13, 14, 15, 16, 17, 18)] ~ gen_vcs_ord[3],
#       gs %in% gen_vcs_list[c(19, 20, 21, 22, 23, 24, 25)] ~ gen_vcs_ord[4],
#       TRUE ~ cs),
#   ) %>%
#   select(subject = Subject,
#          trial = TRIAL_INDEX,
#          time = Time,
#          pupil = Pupil,
#          cs = cs,
#          gs =gs,
#          diss = gs_deg
#          ) %>%
#   mutate(time = time / 1000)
# cat(sprintf("Finished cleaning eye-tracking data for subject %s, task %s, phase %s\n", subject, task, phase))
# ```
#
# #### Pupil dilation in function of GS
#
# ```{r}
# gen_pupil_full <- eye_gen %>%
#   group_by(time, cs) %>%
#   summarize(avg_pupil = mean(pupil, na.rm = TRUE),
#             .groups = "drop")
#
# p_gen_pupil_full <- ggplot(gen_pupil_full, aes(x = time, y = avg_pupil, color = cs)) +
#   geom_line(size = 0.8) +
#   labs(title = "Pupil Dilation Over Time",
#        x = "Time (s)",
#        y = "Baseline-corrected pupil diameter (mm)") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = colors) +
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black")
#   )
#
# # Print the plot
# p_gen_pupil_full
# save_plot(file.path(output_path, "p_gen_pupil_full.png"), p_gen_pupil_full)
#
#
# # average
# gen_avg_win <- gen_pupil_full %>%
#     filter(between(time, window_start, window_end)) %>%
#     group_by(cs) %>%
#     summarize(avg_pupil = mean(avg_pupil, na.rm = TRUE),
#               .groups = "drop")
#
# p_gen_avg_win <- ggplot(gen_avg_win, aes(x = cs, y = avg_pupil, fill = cs)) +
#     geom_bar(stat = "identity", position = "dodge", width = 0.7) +
#     labs(title = "Average Pupil Diameter (0.5s - 1.8s)",
#          x = "Condition",
#          y = "Average baseline-corrected pupil diameter (mm)",
#          fill = "Condition") +
#     scale_fill_manual(values = colors) +
#     theme_minimal()
# p_gen_avg_win
# save_plot(file.path(output_path, "p_gen_avg_win.png"), p_gen_avg_win)
# ```
#
# #### Pupil dilation in function of GS and perceptual dissimilarity
#
# ```{r}
# gen_pupil_diss <- eye_gen %>%
#   group_by(time, cs, diss) %>%
#   summarize(avg_pupil = mean(pupil, na.rm = TRUE))
#
# # Convert 'diss' and 'cs' to factor
# gen_pupil_diss$diss <- as.factor(gen_pupil_diss$diss)
# # Rename to orignal and novel cs
# levels(gen_pupil_diss$diss)[levels(gen_pupil_diss$diss) == "0"] <- "Original"
# levels(gen_pupil_diss$diss)[levels(gen_pupil_diss$diss) == "99"] <- "Novel"
# gen_pupil_diss$cs <- as.factor(gen_pupil_diss$cs)
#
# # Create the plot
# p_gen_pupil_dis <- ggplot(gen_pupil_diss, aes(x = time, y = avg_pupil, color = cs)) +
#   geom_line(size = 0.8) +
#   facet_wrap(~ diss, scales = "free_x", ncol = 4) +
#   labs(title = "Pupil Dilation Over Time by Dissimilarity",
#        x = "Time (s)",
#        y = "Baseline-corrected pupil diameter (mm)") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = colors) +
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black")
#   )
#
# # Print the plot
# p_gen_pupil_dis
# save_plot(file.path(output_path, "p_gen_pupil_dis.png"), p_gen_pupil_dis, width = 18, height = 12)
#
# # visualize avg pupil per dissimilarity
# # note: will want to group by subject, x-axis and color terms, then compute avg
# p_gen_pupilavg_dis <- ggplot(gen_pupil_diss %>%
#                             filter(time >= window_start & time <= window_end),
#                           aes(x = diss, y = avg_pupil, color = cs, group = cs)) +
#   stat_summary(fun = mean, geom = "point") +
#   stat_summary(fun = mean, geom = "line") +
#   labs(title = "Pupil Dilation by Dissimilarity",
#        x = "Dissimilarity",
#        y = "Average baseline-corrected pupil diameter (mm)") +
#   scale_color_manual(values = colors) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black"))
#
# save_plot(file.path(output_path, "p_gen_pupilavg_dis.png"), p_gen_pupilavg_dis)
# # with model fit
# p_gen_pupilavg_dis_fit <- ggplot(gen_pupil_diss %>%
#                             filter(time >= window_start & time <= window_end),
#                           aes(x = diss, y = avg_pupil, color = cs, group = cs, fill = cs)) +
#   geom_smooth(method = "loess") +
#   labs(title = "Pupil Dilation by Dissimilarity",
#        x = "Dissimilarity (degree)",
#        y = "Average baseline-corrected pupil diameter (mm)") +
#   scale_color_manual(values = colors) +
#   scale_fill_manual(values = colors) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black"))
# save_plot(file.path(output_path, "p_gen_pupilavg_dis_fit.png"), p_gen_pupilavg_dis_fit)
# ```
# 
# ## Gaze data
# 
# ```{r}
# gaze_x <- eye_cond_raw %>%
#   select(subject = Subject,
#          trial = TRIAL_INDEX,
#          time = Time,
#          pupil = Pupil,
#          cs = cs,
#          gazex = Gaze_X,
#          ia = RIGHT_INTEREST_AREA_ID) %>%
#   mutate(
#     cs = case_when(
#       cs %in% vcs_list[1] ~ vcs_ord[1],
#       cs %in% vcs_list[2] ~ vcs_ord[2],
#       cs %in% vcs_list[3] ~ vcs_ord[3],
#       TRUE ~ cs
#     )
#   )
# gaze_x <- gaze_x %>%
#   group_by(time, cs) %>%
#     summarize(avg_gazex = mean(gazex, na.rm = TRUE))
# 
# p_cond_gaze <- ggplot(gaze_x, aes(x = time, y = avg_gazex, color = cs)) +
#   geom_line(size = 0.8) +
#   labs(title = "Gaze Direction Over Time",
#        x = "Time (s)",
#        y = "Gaze x") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = colors) +
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black")
#   )
# p_cond_gaze

```

# Backup code

```{r}
# gen_pupil_diss$diss <- as.factor(gen_pupil_diss$diss)
# 
# # Define colors for each level of 'diss'
# diss_colors <- c("Original" = "red", "10" = "blue", "20" = "green", 
#                  "30" = "purple", "40" = "orange", "50" = "pink",
#                  "60" = "brown", "Novel" = "grey")
# 
# # Create the plot
# p_gen_pupil_all <- ggplot(gen_pupil_diss, aes(x = time, y = avg_pupil, color = diss)) +
#   geom_line(size = 0.8) +
#   facet_wrap(~ cs, scales = "free_x", ncol = 4) +
#   labs(title = "Overall Pupil Dilation Over Time",
#         x = "Time (s)",
#         y = "Baseline-corrected pupil diameter (mm)") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = diss_colors) +
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black")
#   )
# 
# # Print the plot
# p_gen_pupil_all


# # overall pupil by congruency
# pupil_full_cong <- cond_cleaned %>%
#   filter(cs != "null") %>%
#   group_by(time, cs, pos_item, incong_loc, iter) %>%
#   summarize(avg_pupil = mean(pupil, na.rm = TRUE), .groups = "drop") %>%
#   mutate(cong = if_else(incong_loc == "FALSE", "congruent", "incongruent"))
# 
# # find null separately
# null_left <- read_csv("/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab/eye/preprocessed/L003_null_left.csv")
# null_right <- read_csv("/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab/eye/preprocessed/L003_null_right.csv")
# 
# null_left <- null_left %>%
#   mutate(ia = "left")
# null_right <- null_right %>%
#   mutate(ia = "right")
# 
# null <- bind_rows(null_left, null_right)
# null <- null %>%
#   select(subject = Subject,
#          trial = TRIAL_INDEX,
#          time = Time,
#          pupil = Pupil,
#          cs = cs,
#          ia = ia) %>%
#   mutate(
#     cstype = case_when(
#       cs == "VCS_10.jpg" ~ "null"),
#     cs = cstype,
#     time = time / 1000)
# 
# pupil_null_cong <- null %>%
#   mutate(cong = if_else(ia == "left", "congruent", "incongruent")) %>%
#   group_by(time, cong, cs, trial) %>%
#   summarize(avg_pupil = mean(pupil, na.rm = TRUE), .groups = "drop")
# 
# # merge all CS types
# us_time_cong <- bind_rows(pupil_full_cong, pupil_null_cong)
# 
# 
# # Create interaction term for the legend
# us_time_cong$Congruency <- interaction(us_time_cong$cs, us_time_cong$cong)

# Plot with separate graphs for each congruency
# p_pupil_full_cong <- ggplot(us_time_cong, aes(x = time, y = avg_pupil, color = cs)) + # Change here
#   geom_line(size = 0.8) +
#   facet_wrap(~ cong, scales = "free_y") + # Facet by congruency
#   labs(title = "Pupil Dilation Over Time in Congruent ROI",
#        x = "Time after US onset (s)",
#        y = "Baseline-corrected pupil diameter (mm)") +
#   scale_x_continuous(limits = c(0, 3), breaks = breaks_x, labels = labels_x) +
#   scale_y_continuous(breaks = breaks_y) +
#   scale_color_manual(values = colors) + 
#   geom_vline(xintercept = window_start) +
#   geom_vline(xintercept = window_end) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color = "black"))
# save_plot(file.path(output_path, "p_pupil_full_cong.png"), p_pupil_full_cong)

# # window average by congruency
# pupil_avg_cong <- us_time_cong %>%
#   filter(between(time, window_start, window_end)) %>%
#   group_by(cs, cong) %>%
#   summarize(avg_pupil = mean(avg_pupil, na.rm = TRUE)) %>%
#   ungroup()
# pupil_avg_cong
# 
# # Define colors
# intercolors <- c("alc.congruent" = "#e4b2db", "neut.congruent" = "#73cebc", "null.congruent" = "#0773a9",
#             "alc.incongruent" = "#cc99c2", "neut.incongruent" = "#66c2a4", "null.incongruent" = "#055a86")
# pupil_avg_cong$cong <- factor(pupil_avg_cong$cong, levels = c("congruent", "incongruent"))
# pupil_avg_cong$Congruency <- interaction(pupil_avg_cong$cs, pupil_avg_cong$cong)
# legend_lab <- c("alc.congruent" = "congruent",
#                 "alc.incongruent" = "incongruent",
#                 "neut.congruent" = "congruent",
#                 "neut.incongruent" ="incongruent",
#                 "null.congruent" = "congruent",
#                 "null.incongruent" = "incongruent")
# 
# # Convert "cong" to a factor and create an interaction term for coloring
# pupil_avg_cong$cong <- factor(pupil_avg_cong$cong, levels = c("congruent", "incongruent"))
# pupil_avg_cong$Congruency <- interaction(pupil_avg_cong$cs, pupil_avg_cong$cong)
# 
# p_avg_pupil_cong <- ggplot(pupil_avg_cong, aes(x = cs, y = avg_pupil, fill = Congruency)) +
#   geom_bar(stat = "identity", position = "dodge", width = 0.7) +
#   labs(title = "Average Pupil Diameter by Congruency (0.5s - 1.8s)",
#        x = "Condition",
#        y = "Average baseline-corrected pupil diameter (mm)") +
#   scale_fill_manual(values = intercolors, labels = legend_lab) +
#   theme_minimal() +
#   theme(plot.title.position = "plot",
#         plot.title = element_text(hjust = 0.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(color="black"))
# save_plot(file.path(output_path, "p_avg_pupil_cong.png"), p_avg_pupil_cong)
# ```
# 
# #### Quartile \~ Congruency
# 
# ```{r}
#   # recode
#   pupil_null_cong <- pupil_null_cong %>%
#     mutate(quartile = ntile(iter, 4)) %>%
#     ungroup()
# 
#   pupil_csneut_cong <- pupil_full_cong %>%
#     mutate(quartile = ntile(iter, 4)) %>%
#     ungroup() %>%
#     select(time, cong, cs, avg_pupil, iter, quartile)
#   
#   pupil_null_cong$trial <- NULL
#   allcs_cong <- bind_rows(pupil_null_cong, pupil_csneut_cong)
#   
#   # calculate the average pupil dilation for each quartile for each condition
#   
#   avg_cong_quart <- allcs_cong %>%
#     filter(between(time, window_start, window_end)) %>%
#     group_by(cs, quartile, cong) %>%
#     summarize(avg_pupil = mean(avg_pupil, na.rm = TRUE),
#               .groups = "drop")
#     
#   # bar graph for average pupil dilation between 0.5~1.8s
#   p_avg_cong_quart <- ggplot(avg_cong_quart, aes(x = quartile, y = avg_pupil, fill = cs)) +
#     geom_bar(stat = "identity", position = "dodge", width = 0.7) +
#     facet_wrap(~cong, scales = "free_y") +
#     labs(title = "Average Pupil Diameter by Quartile and Congruency",
#          x = "Quartile",
#          y = "Average baseline-corrected pupil diameter (mm)",
#          fill = "Condition") +
#     scale_fill_manual(values = colors) +
#     theme_minimal() +
#     theme(plot.title.position = "plot",
#           plot.title = element_text(hjust = 0.5),
#           panel.grid.major = element_blank(),
#           panel.grid.minor = element_blank(),
#           axis.line = element_line(color="black"))
# 
#   print(p_avg_cong_quart)
#   save_plot(file.path(output_path, "p_avg_cong_quart.png"), p_avg_cong_quart)
#   
# 
#   
# # divide into iterations
#   pupil_iter <- allcs_cong %>%
#     group_by(cs, iter, cong) %>%
#     filter(time >= window_start & time <= window_end) %>%
#     summarize(avg_pupil = mean(avg_pupil, na.rm = T))
# 
#   avg_cong_quart$quartile <- factor(avg_cong_quart$quartile,
#                                      levels = 1:4,
#                                      labels = c("Q1: Repetition 1~10", "Q2: Repetition 11~20", "Q3: Repetition 21~30", "Q4: Repetition 31~40"))
# 

```
