---
title: "Step3_preproc"
author: "zihan bai"
date: "2023-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# Load necessary packages
```{r }
# install.packages("pacman)
library(pacman)
p_load(readr, dplyr, ggplot2, PupillometryR, PupilPre, shiny, tidyverse, stringr, emmeans, sjPlot, nlme, patchwork, purrr, ggpubr, lme4,PMCMRplus, stats, zoo, splines, psych)
```
# Paths
```{r}
# If work on Mac
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"

# If work on Windows
# base_path <- "C:/Users/Zihan/Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"

# Input
preproc_path <- file.path(base_path, "pupillometry", "preprocessed")
# preprocessed_cond_cs_path <- file.path(preproc_path, 'pupil_size', paste0('preprocessed_cond_cs.RData'))
# # Load existing preprocessed data if it exists
# if (file.exists(preprocessed_cond_cs_path)) {
#   load(preprocessed_cond_cs_path)
# }

preprocessed_cond_cs_all_path <- file.path(preproc_path, 'pupil_size', paste0('preprocessed_cond_cs_all.RData'))
# Load existing preprocessed data if it exists
if (file.exists(preprocessed_cond_cs_all_path)) {
  load(preprocessed_cond_cs_all_path)
}

# Which subjects?
subjects <- c('L003', 'L004', 'L006', 'L007', 'L009', 'L020', 'L021', 'H001', 'L022', 'L024', 'L026', 'L028', 'L029', 'L013', 'L014', 'L015', 'L017', 'L019', 'H002', 'H003', 'L023', 'L031')

n_distinct(subjects)
n_distinct(preprocessed_cond_cs_all$subject)

excl_subjects <- c('L025')
```
# Conditioning
## CS onset
### Load behavior data
```{r}
# Load behavioral data
behav_fname <- paste0(base_path, "/", "processed_behav", "/", "combined_conditioning_data.csv")
behav_cond <- read_csv(behav_fname)
recode_behav_cond <- behav_cond %>%
  rename(trial = trial_num,
         subject = subId) %>%
  mutate(incong_loc = case_when(
          trial_cat %in% c("alc_null_surprise", "null_alc_surprise", 
                          "null_neut_surprise", "neut_null_surprise") ~ 'NOTHING',
          # ignore 30% trials while congruent locs contain nothing
          # so each cs is reinforced with the same us location in 70% of trials
          # and with the same us image in 85% of trials
          stim == 'null' ~ 'NOTHING',
          TRUE ~ as.character(incong_loc))) %>%
  filter(subject %in% subjects) %>%
  select(c(subject, phase, trial, incong_loc, stim, iter, drink_2grp))

n_distinct(recode_behav_cond$subject)
```
### Combine behavior and pupil
```{r}
combined_cond_cs <- left_join(preprocessed_cond_cs_all, recode_behav_cond, by = c("subject", "trial", "stim"))

combined_cond_cs$subject <- as.factor(combined_cond_cs$subject)
combined_cond_cs$time <- as.numeric(as.character(combined_cond_cs$time))

# Quick check to ensure row counts are as expected
nrow(combined_cond_cs)
nrow(preprocessed_cond_cs_all)
```

### Pupil response after cs onset by stimulus-type
```{r} 
# For an aggregate view across subjects to highlight general trends rather than individual variations, we calcualte the mean size at each time point for each condition. This provides a representation of how the average pupil dilation changes over time in response to each stimulus type.

# Time-course
# trace_plot_cond_cs <- combined_cond_cs %>%
#   filter(stim %in% c("alc", "neut", "null")) %>%
#   group_by(drink_2grp, time, stim) %>%
#   summarise(avg_ps = mean(ps_cor, na.rm = TRUE), .groups = 'drop')

# check saccades frequency during window by stim

P_trace_plot_cs <- ggplot(combined_cond_cs, aes(x = time, y = ps_cor, color = stim)) +
  facet_grid(~ drink_2grp) +
  # geom_line() +
  geom_smooth(se=FALSE) +
  labs(title = "Pupil dilation in response to CS presentation",
       x = "Time (s)",
       y = "Pupul dilation",
       color = "US") + 
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = 1.8, linetype = "dashed", color = "grey") +
  scale_color_manual(values = c("alc" = "#A06593", "neut" = "#6086C5", "null" = "grey")) +
  theme_classic()

# compare pre vs post incongruent trials
dat_per_stim <- combined_cond_cs %>% select(c("subject", "stim", "trial", "incong_loc")) %>% distinct()

# recode order per stimulus type
dat_per_stim <- dat_per_stim %>% group_by(subject, stim) %>% mutate(stimrep = row_number())

# flag pre/post incong
incong_loc_trials <- dat_per_stim %>% filter(incong_loc=="TRUE") %>%
  mutate(prerep = stimrep-1,
         postrep = stimrep+1)

incong_outcome_trials <- dat_per_stim %>% filter(incong_loc=="NOTHING") %>%
  mutate(prerep = stimrep-1,
         postrep = stimrep+1)

# add info to dat_per_stim df
for (s in sort(unique(dat_per_stim$subject))) {
  dat_sub <- dat_per_stim %>% filter(subject==s)
  incong_loc_sub <- incong_loc_trials %>% filter(subject==s)
  incong_outcome_sub <- incong_outcome_trials %>% filter(subject==s)
  
  dat_sub$pre_incong_loc = 0
  dat_sub$post_incong_loc = 0
  
  dat_sub$pre_incong_outcome = 0
  dat_sub$post_incong_outcome = 0
  
  
  for (stimname in sort(unique(dat_per_stim$stim))) {
    dat_stim <- dat_sub %>% filter(stim==stimname)
    # location of US
    stim_pre_idx <- which(dat_stim$stimrep %in% incong_loc_sub[incong_loc_sub$stim==stimname,]$prerep)
    dat_stim[stim_pre_idx,]$pre_incong_loc = 1
    
    stim_post_idx <- which(dat_stim$stimrep %in% incong_loc_sub[incong_loc_sub$stim==stimname,]$postrep)
    dat_stim[stim_post_idx,]$post_incong_loc = 1
    
    # US type
    stim_pre_idx <- which(dat_stim$stimrep %in% incong_outcome_sub[incong_outcome_sub$stim==stimname,]$prerep)
    dat_stim[stim_pre_idx,]$pre_incong_outcome = 1
    
    stim_post_idx <- which(dat_stim$stimrep %in% incong_outcome_sub[incong_outcome_sub$stim==stimname,]$postrep)
    dat_stim[stim_post_idx,]$post_incong_outcome = 1
    
    if (stimname=='alc') {
    dat_stim_concat <- dat_stim
    } else {
      data_stim_concat <- rbind(dat_stim_concat, dat_stim)
    }
  } # stim loop
  
  if (s=='H001') {
    dat_sub_concat <- data_stim_concat
  } else {
    dat_sub_concat <- rbind(dat_sub_concat, data_stim_concat)
  }
} # subject loop

# plot pre- and post (where are the other stim...)
ggplot(combined_cond_cs %>% merge(dat_sub_concat, by = c("subject", "stim", "trial", "incong_loc")) %>%
         mutate(incong_loc_timing = ifelse(pre_incong_loc==1, "pre", ifelse(post_incong_loc==1, "post", NA))) %>% filter(!(is.na(incong_loc_timing))), 
       aes(x = time, y = ps_cor, color = incong_loc_timing)) +
  facet_grid( stim~ drink_2grp) +
  # geom_line() +
  geom_smooth(se=FALSE) +
  labs(title = "Pupil dilation in response to CS presentation: Pre/post incongruent locations",
       x = "Time (s)",
       y = "Pupul dilation",
       color = "US") + 
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = 1.8, linetype = "dashed", color = "grey") +
  # scale_color_manual(values = c("alc" = "#A06593", "neut" = "#6086C5", "null" = "grey")) +
  theme_classic()

ggplot(combined_cond_cs %>% merge(dat_sub_concat, by = c("subject", "stim", "trial", "incong_loc")) %>%
         mutate(incong_outcome_timing = ifelse(pre_incong_outcome==1, "pre", ifelse(post_incong_outcome==1, "post", NA))) %>% filter(!(is.na(incong_outcome_timing))), 
       aes(x = time, y = ps_cor, color = incong_outcome_timing)) +
  facet_grid( stim~ drink_2grp) +
  # geom_line() +
  geom_smooth(se=FALSE) +
  labs(title = "Pupil dilation in response to CS presentation: Pre/post incongruent outcomes",
       x = "Time (s)",
       y = "Pupul dilation",
       color = "US") + 
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = 1.8, linetype = "dashed", color = "grey") +
  # scale_color_manual(values = c("alc" = "#A06593", "neut" = "#6086C5", "null" = "grey")) +
  theme_classic()
```

```{r}
# Save the figure
pdf(file = 'P_trace_plot_cs.pdf', width = 8, height = 4, useDingbats = F)
print(P_trace_plot_cs)
dev.off()
```


### Window analysis
```{r}
# In contrast, we calculate the mean pupil changes for each condition for each subject and controls for the individual differences among subjects
start_window <- 0.5
end_window <- 1.8
# end_window <- 1

win_cond_cs <- combined_cond_cs %>%
  filter(time >= start_window & time <= end_window,
         stim %in% c("alc", "neut", "null")) %>%
  group_by(subject, stim, drink_2grp, trial) %>%
  summarize(win_avg = mean(ps_cor, na.rm = TRUE), .groups = 'drop')

P_win_cs <- ggplot(win_cond_cs, aes(x = stim, y = win_avg, fill = stim)) +
  facet_grid(~ drink_2grp) +
  geom_violin() +
  labs(x = "US",
       y = "Mean pupil dilation",
       fill = "US") + 
  scale_fill_manual(values = c("alc" = "#A06593", "neut" = "#6086C5", "null" = "grey")) +
  theme_classic()
print(P_win_cs)
```

```{r}
# Save the figure
pdf(file = 'P_win_cs.pdf', width = 4, height = 4, useDingbats = F)
print(P_win_cs)
dev.off()
```

### Within-subject pupil changes by condition
```{r}
subject_condition_averages <- win_cond_cs %>%
  group_by(drink_2grp, subject, stim) %>%
  summarise(mean_ps_cor = mean(win_avg, na.rm = TRUE), .groups = 'drop')

P_win_ind_cs <- ggplot(subject_condition_averages, aes(x = stim, y = mean_ps_cor, group = subject)) +
  facet_grid(~ drink_2grp) +
  geom_line(aes(color = subject), show.legend = FALSE) + # Connects points with lines for each subject
  geom_point(aes(color = subject)) + # Plots the actual points
  labs(title = "Within-subject pupil size across conditions",
       x = "CS",
       y = "Mean pupil size") +
  theme_minimal() +
  scale_color_viridis_d() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(P_win_ind_cs)
```
```{r}
# Save the figure
pdf(file = 'P_win_ind_cs.pdf', width = 4, height = 4, useDingbats = F)
print(P_win_ind_cs)
dev.off()
```

### Split into halves
```{r}
group_avg_cond_cs_split <- combined_cond_cs %>%
  mutate(split = case_when(iter<=20 ~ "firstH", iter>20 ~ "secondH")) %>%
  group_by(split, time, stim, drink_2grp) %>%
  summarise(avg_ps = mean(ps_cor, na.rm = TRUE),
            se = sd(ps_cor, na.rm = TRUE) / sqrt(n()),
            .groups = 'drop')
P_trace_plot_cs_split <- ggplot(group_avg_cond_cs_split, aes(x = time, y = avg_ps, color = stim, group = interaction(stim, drink_2grp))) +
  geom_line() +
  facet_grid(drink_2grp ~ split) + # Adjusted for a grid by drinker group and split
  labs(title = "Pupil dilation in response to CS presentation",
       x = "Time (s)",
       y = "Pupil dilation", # Corrected typo
       color = "Stimulus Type") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = 1.8, linetype = "dashed", color = "grey") +
  scale_color_manual(values = c("alc" = "#A06593", "neut" = "#6086C5", "null" = "grey")) +
  theme_classic()
print(P_trace_plot_cs_split)
```

```{r}
# Save the figure
pdf(file = 'P_trace_plot_cs_split.pdf', width = 4, height = 4, useDingbats = F)
print(P_trace_plot_cs_split)
dev.off()
```

```{r}
win_cond_cs_split <- combined_cond_cs %>%
  filter(time >= start_window & time <= end_window) %>%
  mutate(split = case_when(iter<=20 ~ "firstH", iter>20 ~ "secondH")) %>%
  group_by(split,subject, stim, drink_2grp) %>%
  summarize(win_avg = mean(ps_cor, na.rm = TRUE), .groups = 'drop')

P_win_cond_cs_split <- ggplot(win_cond_cs_split, aes(x = stim, y = win_avg, fill = split)) +
  facet_grid(~ drink_2grp) +
  geom_boxplot() +
  labs(x = "US",
       y = "Pupil dilation",
       color = "US") + 
  theme_classic()
print(P_win_cond_cs_split)
```
```{r}
# Save the figure
pdf(file = 'P_win_cond_cs_split.pdf', width = 4, height = 4, useDingbats = F)
print(P_win_cond_cs_split)
dev.off()
```

### Statistical Analysis
```{r}
# the effects of stimulus type, drinking level, and repetition on pupil dilation.
model_aov <- aov(mean_ps ~ stim, data = group_avg_cond_cs)
summary(model_aov)

# Model pupil size as a function of stimulus type, drinking level, and repetition, including random effects for subject
model <- lmer(ps ~ stim + (1 | subject), data = combined_cond_cs)

summary(model)
```
### Repetition
``` {r}
# divide into iterations
cond_cs_iter <- combined_cond_cs %>%
  filter(stim %in% c("alc", "neut", "null"),
         subject!="H001") %>%
  group_by(subject, stim, iter) %>%
  summarize(avg_ps = mean(ps_cor, na.rm = TRUE)) %>%
  ungroup()


# lm
lm_cond_cs_iter <- lme(avg_ps ~ stim * iter, random=~1|subject, data = cond_cs_iter)
anova_results <- anova(lm_cond_cs)
print(anova_results)
  
plot_model(lm_cond_cs, type = "eff", terms = c("iter", "stim"), axis.title = c("Repetition", "Pupil dilation")) +
  scale_color_manual(values = c("alc" = "#A06593", "neut" = "#6086C5", "null" = "grey")) +
  theme_classic()
```


## Towardness
```{r}
# Which subjects?
subjects <- c('L003', 'L004', 'L006', 'L007', 'L009', 'L020', 'L021', 'H001', 'L022', 'L024', 'L026', 'L027', 'L028', 'L029')

excl_subjects <- c('L025')

# In what sample rate?
s_rate_500 <- c('L003', 'L004', 'L006', 'L007', 'L009', 'L020', 'L021', 'H001')
s_rate_250 <- c('L013', 'L014', 'L015', 'L017', 'L019', 'H002', 'H003')
s_rate_1000 <- c('L025', 'L031')

# Define paths
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"
```

```{r}
# Load anticipation data
append_cond_gaze_path <- file.path(preproc_path, "gaze", "append_cond_gaze.RData")
append_cond_gaze <- list()
if (file.exists(append_cond_gaze_path)) {
  load(append_cond_gaze_path)
}

for (subject in subjects) {
  input_path <- file.path(trial_split_path, subject, paste0(subject, "_cond_split.RData"))
  data <- load(input_path)
  ant <- subject_phases_list$ant_onset %>%
    filter(!is.na(incong_loc)) %>%
    rename(trial = trial_num) %>%
    group_by(trial) %>%
    mutate(time = (row_number() - 1) / 1000 * 2) %>%
    ungroup()
  
  append_cond_gaze[[subject]] <- ant
}
save(append_cond_gaze, file = append_cond_gaze_path)

append_cond_gaze_df <- bind_rows(append_cond_gaze)

for (subject in subjects) {
  input_path <- file.path(trial_split_path, subject, paste0(subject, "_cond_split.RData"))
  data <- load(input_path)
  fix <- subject_phases_list$fixation %>%
    filter(!is.na(incong_loc)) %>%
    rename(trial = trial_num) %>%
    group_by(trial) %>%
    mutate(time = (row_number() - 1) / 1000 * 2) %>%
    ungroup()
  
  append_cond_fix[[subject]] <- fix
}
append_cond_fix_df <- bind_rows(append_cond_fix)
```


``` {r}
gaussian_kernel <- function(sd, len = 30) {
  seq <- seq(-len/2, len/2, length.out = len)
  kernel <- dnorm(seq, mean = 0, sd = sd)
  kernel / sum(kernel)
}
sd_points <- 5  # Standard deviation in terms of data points
kernel <- gaussian_kernel(sd = sd_points)
```


### Method 2
```{r}
# Before averaging across participants, gaze time courses were smoothed by a Gaussian kernel with a 10 ms s.d. (Ede et al., 2019)

# Linear Interpolation on the raw data to fill in missing values.
# Gaussian Kernel Smoothing of the interpolated data to reduce noise.
# Normalization of the smoothed data using a baseline period to control for individual differences.
# Calculation of towardness from the smoothed and normalized data.

Sdata <- make_pupillometryr_data(data = append_cond_gaze_df, subject = subject, trial = trial, time = time, condition = stim)

missing <- calculate_missing_data(Sdata, pupil = xp)
excl_data <- clean_missing_data(Sdata,
                                pupil = xp,
                                trial_threshold = .75,
                                subject_trial_threshold =.75)
int_data <- interpolate_data(data = excl_data, pupil = xp, type = "linear")
  

smooth_data <- int_data %>%
  group_by(subject, trial) %>%
  mutate(xp = ifelse(is.na(xp), NA, 
                                          stats::filter(xp, kernel, sides = 2))) %>%
  ungroup()

fix_median <- append_cond_fix_df %>%
  group_by(subject, trial) %>%
  summarize(fix_median = median(xp, na.rm = TRUE)) %>%
  ungroup()

fix_corr <- smooth_data %>%
  left_join(fix_median, by = c("subject", "trial"))
  
fix_norm <- fix_corr %>%
  group_by(subject, trial) %>%
    # figure out expected position of US
  mutate(expect_loc = ifelse(stim == "null", "none",
                               ifelse(incong_loc == "FALSE", us_loc,
                                      ifelse(us_loc=="right", "left",
                                             ifelse(us_loc=="left", "right", NA))))) %>%
  mutate(xp_norm.fix = (xp - fix_median) / fix_median, # positive = right, negative = left
         towardness.fix = ifelse(expect_loc=="left", -1 * xp_norm.fix, 
                               ifelse(expect_loc=="right", xp_norm.fix,
                                           ifelse(expect_loc == "none", NA, NA)))) %>%
  ungroup()

# smooth_fix_norm <- fix_norm %>%
#   group_by(subject, trial) %>%
#   mutate(smoothed_towardness.fix = ifelse(is.na(towardness.fix), NA, 
#                                           stats::filter(towardness.fix, kernel, sides = 2))) %>%
#   ungroup()
         
```

```{r}
preprocessed_towardness <- file.path(preproc_path, "gaze", "preprocessed_cond_towardness.RData")
load(preprocessed_towardness)

P_cond_towardness <- ggplot(fix_norm %>% filter(stim != "null"), aes(x = time, y = towardness.fix, color = stim)) + geom_smooth(se = TRUE) + labs(y="Towardness", x="Time") + scale_color_manual(values = c("alc" = "#A06593", "neut" = "#6086C5")) + theme_classic() + geom_hline(yintercept = 0)

P_cond_sidebias <- ggplot(fix_norm %>% filter(stim == "null"), aes(x = time, y = xp_norm.fix, color = stim)) + geom_smooth(se = FALSE) + labs(y="Towardness", x="Time") + scale_color_manual(values = c("null" = "grey")) + theme_classic() + geom_hline(yintercept = 0)

```

```{r}
# Save the figure
pdf(file = 'P_cond_towardness.pdf', width = 4, height = 4, useDingbats = F)
print(P_cond_towardness)
dev.off()

pdf(file = 'P_cond_sidebias.pdf', width = 4, height = 4, useDingbats = F)
print(P_cond_sidebias)
dev.off()

# Save the file
output_towardness_path <- file.path(base_path, "pupillometry", "preprocessed", "gaze", "preprocessed_cond_towardness.RData")
save(fix_norm, file = output_towardness_path)
```

### Method 1
```{r}
ant_median <- append_cond_gaze_df %>%
    group_by(subject, trial) %>%
    summarize(ant_median = median(xp, na.rm = TRUE)) 

ggplot(ant_median, aes(x=trial, y=ant_median, group=subject))+ geom_point(aes(color=subject)) + geom_smooth(aes(color=subject), method="loess", se=FALSE) + scale_color_viridis_d() + theme_classic()

ant_corr <- append_cond_gaze_df %>%
  left_join(ant_median, by = c("subject", "trial"))

ant_norm <- ant_corr %>%
  group_by(subject, trial) %>%
    # figure out expected position of US
  mutate(expect_loc = ifelse(stim == "null", "none",
                               ifelse(incong_loc == "FALSE", us_loc,
                                      ifelse(us_loc=="right", "left",
                                             ifelse(us_loc=="left", "right", NA))))) %>%
  mutate(xp_norm.ant = (xp - ant_median) / ant_median,   # positive = right, negative = left
         towardness.ant = ifelse(expect_loc=="left", -1 * xp_norm.ant, 
                               ifelse(expect_loc=="right", xp_norm.ant,
                                           ifelse(expect_loc == "none", NA, NA))))

sanity_check_twd <- ant_norm %>%
  group_by(subject, trial) %>%
  summarize(pos = sum(xp_norm.ant > 0, na.rm = TRUE),
            neg = sum(xp_norm.ant < 0, na.rm = TRUE),
            counts_equal = pos == neg,
            deviation = ifelse(counts_equal, 0, abs(pos - neg))) %>%
  ungroup()

# By normalzing the gaze position by the median gaze position during anticipation (rather than fixation), the number of positive and negative values of normalized gaze position has to be the same, or nearly the same
ggplot(sanity_check_twd, aes(x = deviation)) + geom_histogram(binwidth = 1, fill = "#6086C5", color = "black") + labs(x = "Deviation", y = "Frequency") + theme_classic()
ggplot(sanity_check_twd, aes(x = trial, y = deviation, color = subject)) + geom_point() + geom_smooth(se = FALSE) + labs(x = "Trial", y = "Deviation") + theme_classic()
```

```{r }
# Visualize
ggplot(ant_norm %>% filter(stim != "null"), aes(x = time, y = towardness.ant, color = stim)) +
  geom_smooth(se = FALSE) + theme_classic() + geom_hline(yintercept = 0)

# Within-subjects?
ggplot(twd_cond_all_exclusion %>% filter(stim != "null"), aes(x = time, y = xp_norm.ant, color = expect_loc)) +
  facet_wrap(~subject) +
  geom_smooth(se = FALSE) + theme_classic() + geom_hline(yintercept = 0)

# side bias?
ggplot(twd_cond_all_exclusion %>% filter(stim == "null"), aes(x = time, y = xp_norm.ant, color = stim)) + geom_smooth(se = FALSE) + theme_classic() + geom_hline(yintercept = 0)
  
```


```{r}
# look for reaction time for alc and neut because the time they take to move to the expected location seems different?
```


## US onset (Secondary)
```{r}

```

### Pupil response after us onset by stimulus-type
```{r}
# Time-course
ggplot(merged_us_df, aes(x = time, y = ps_cor, group = stim, color = stim)) +
  stat_summary(fun = mean, geom = "line", na.rm = TRUE) +
  theme_minimal()

# Window analyses
# Window of interest (0.5-1.8s)
window_start = 0.5
window_end = 1.8
combined_us_window <- filter(merged_us_df, between(time, window_start, window_end))
us_window <- create_window_data(combined_us_window, pupil = ps_cor)
plot(us_window, pupil = ps_cor, windows = F, geom = 'boxplot')

```
# Generalization
## GS onset
### Load behavior and pupil
```{r}
# Which subjects?
subjects <- c('L003', 'L004', 'L006', 'L007', 'L009', 'L020', 'L021', 'H001', 'L022', 'L024', 'L026', 'L027', 'L028', 'L029')

excl_subjects <- c('L025')

# Load pupil from Step 3
preprocessed_gen_gs_path <- file.path(preproc_path, 'pupil_size', paste0('preprocessed_gen_gs.RData'))
# Load existing preprocessed data if it exists
if (file.exists(preprocessed_gen_gs_path)) {
  load(preprocessed_gen_gs_path)
}

# Load behavior in generalization task
behav_fname <- paste0(base_path, "/", "processed_behav", "/", "combined_generalization_data.csv")

behav_gen <- read_csv(behav_fname) %>%
  filter(block != 'prac') %>%
  rename(subject = subId,
         diss = gs_deg,
         stim = trial_cat) %>%
  group_by(subject) %>%
  mutate(trial = row_number()) %>%
  ungroup() %>%
  filter(subject %in% subjects) %>%
  select(subject, phase, trial, stim, diss, drink_2grp)
```

### Combine behavior and pupil
```{r}
combined_gen_gs <- left_join(processed_data, behav_gen, by = c("subject", "trial", "stim"))

combined_gen_gs$subject <- as.factor(combined_gen_gs$subject)
combined_gen_gs$time <- as.numeric(as.character(combined_gen_gs$time))
# Quick check to ensure row counts are as expected
nrow(combined_gen_gs)
nrow(processed_data)
``` 

### Pupil response after gs onset by stimulus-type
```{r}
# Card-evoked pupil dilation will be analyzed with a linear mixed effects model with degree of similarity to original CS, US category, drinking level, and age as predictors, with time spent doing the task as a covariate of no interest.
P_traceplot_gen_gs <- ggplot(combined_gen_gs %>% filter(stim!="novel", subject!="H001"), aes(x=time, y=ps_cor, color=as.factor(diss))) + 
  geom_smooth() + 
  facet_grid(~ stim, scales = "free_y") + 
  labs(x = "Time", y = "Pupil dilation", color = "Dissimilarity") +
  theme_minimal()

# alc_cs = list(deg0 = rnorm(100), deg10 = rnorm(100), deg20 = rnorm(100))
# neut_cs = list(deg0 = rnorm(100), deg10 = rnorm(100), deg20 = rnorm(100))
# 
# alc_cs.df <- alc_cs %>% as.data.frame() %>% mutate(cs_type = "alc", timing = c(1:100))
# neut_cs.df <- neut_cs %>% as.data.frame() %>% mutate(cs_type = "neut", timing = c(1:100))
# 
# mult_cs.df <- rbind(alc_cs.df, neut_cs.df)
# 
# mult_cs.df %>% 
#   pivot_longer(c(1:3), values_to = "pupil_area", names_to = "deg") %>%
#   ggplot(aes(x = timing, y = pupil_area, color = deg)) +
#   geom_line() +
#   facet_grid(~cs_type) +
#   theme_classic()

ggplot(combined_gen_gs %>% filter(stim!="novel", subject!="H001"), aes(x=time, y=ps_cor, color=stim)) + 
     geom_smooth() + 
     facet_grid(~ as.factor(diss), scales = "free_y") + 
     labs(x = "Time", y = "Pupil dilation", color = "Dissimilarity") +
     theme_minimal()
```

```{r}
# Save the figure
pdf(file = 'P_traceplot_gen_gs.pdf', width = 8, height = 4, useDingbats = F)
print(P_traceplot_gen_gs)
dev.off()
```
