---
title: "Step1_asc_split"
author: "zihan bai"
date: "2023-12-18"
output: html_document
---
Step1: this script takes asc files converted from edf for each subject in , and splits them into 4 individuals files for different tasks.
e.g. In: L003.asc -- Out: L003_cond_csv, L003_gen_csv, L003_csmem_csv, L003_usmem_csv

# Setup
```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(eyelinker, intervals, stringr, readr, dplyr, ggplot2, PupillometryR, PupilPre, shiny, tidyverse, stringr, emmeans, sjPlot, nlme, patchwork, purrr, ggpubr, lme4, psych)
```

# Split into subject/tasks
## v1

------------------------------------------------------------------------------------------------
Use this version if the subject has 1 giant EDF and contains 99 trials in gen task instead of 100
```{r}
# Function to find closest time
# Sometimes the time when a specific msg appears does not match with the time in raw samples, therefore we define a function to find he close time and merge raw and msg without losing any msgs
find_closest_time <- function(time, df) {
  abs_diff <- abs(df$time - time)
  closest_time <- df$time[which.min(abs_diff)]
  return(closest_time)
}

# Define subjects and base paths
subjects <- c('H002', 'H003', 'L013', 'L014', 'L015', 'L017', 'L019', 'L023')

# Run on Mac:
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"

# Run on Windows
#base_path <- "C:/Users/Zihan/Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"

asc_split_path <- file.path(base_path, "pupillometry", "restructured", "asc_split")
if (!dir.exists(asc_split_path)) {
  dir.create(asc_split_path)
}

# Define task phases
task_phases <- list(
  cond = c(41, 160),
  gen = c(167, 265), # for v1, use 167
  csmem = c(266, 301),
  usmem = c(302, 409)
)

# Initialize an empty list to store data for all subjects
all_data <- list()

# Read and process data for each subject
for (subject in subjects) {
  input_path <- file.path(base_path, "pupillometry", "raw_asc", sprintf("%s.asc", subject))
  df <- read.asc(input_path)
  
  print(paste0('Subject ', subject, ' sample rate: ', df$info$sample.rate))
  
  df$msg$closest_time <- sapply(df$msg$time, find_closest_time, df = df$raw)
  df$raw$closest_time <- df$raw$time # For merging
  
  combined_df <- merge(df$raw, df$msg, by = c('block', 'closest_time'), all.x = TRUE)
  
  combined_df$time <- combined_df$closest_time
  combined_df$subject <- subject
  combined_df <- select(combined_df, c('subject','block', 'time', 'xp', 'yp', 'ps', 'text'))
  
  
  # Store each task phase for the current subject
  df_phase <- list()
  for (task_name in names(task_phases)) {
    phase_range <- task_phases[[task_name]]
    start_trial <- phase_range[1]
    end_trial <- phase_range[2]
    
    # Filter the current phase range
    df_filtered <- combined_df %>%
      filter(between(block, start_trial, end_trial))
    
    print(paste0('Subject ', subject, ': Trials found in ', task_name, ': ', n_distinct(unique(df_filtered$block))))
    
    # Store the filtered dataframe in df_phase with the phase name as the key
    df_phase[[task_name]] <- df_filtered
  }
  
  # Store the task phase data for the current subject in all_data
  all_data[[subject]] <- df_phase
  
  subject_folder <- file.path(asc_split_path, subject)
  if (!dir.exists(subject_folder)) {
    dir.create(subject_folder)
  }
  
  for (task_name in names(task_phases)) {
    # DataFrame for the current subject and phase
    df_to_save <- all_data[[subject]][[task_name]]
    
    # Construct the filename
    file_name <- paste0(subject, "_", task_name, ".RData")
    file_path <- file.path(subject_folder, file_name)
    
    # Save
    save(df_to_save, file = file_path)
  }
  print('Completed!')
}
```


## v2

------------------------------------------------------------------------------------------------
Use this version if the subject already has interim EDFs and contains 99 trials in gen task instead of 100
```{r}
rm(list = ls())

# Function to find closest time
# Sometimes the time when a specific msg appears does not match with the time in raw samples, therefore we define a function to find he close time and merge raw and msg without losing any msgs
find_closest_time <- function(time, df) {
  abs_diff <- abs(df$time - time)
  closest_time <- df$time[which.min(abs_diff)]
  return(closest_time)
}

# Define subjects and base paths
subjects <- c('L025')

# Run on Mac:
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"

asc_split_path <- file.path(base_path, "pupillometry", "restructured", "asc_split")
if (!dir.exists(asc_split_path)) {
  dir.create(asc_split_path)
}


# Create a list of the different blocks the EDFs were originally split into
blocks <- c('c1','c2','cm','g1','g2','u1','u2','se')

# Create a list of the task phases as we want to name the output files
task_phases <- c('cond','csmem','gen','usmem','fixation')

# Read and process data for each subject
for (subject in subjects) {
  
  # Create a folder in "asc_split" where output will go
  subject_folder <- file.path(asc_split_path, subject)
  if (!dir.exists(subject_folder)) {
    dir.create(subject_folder)
  }
  
  # Initiate an empty list to store all blocks for each subject
  block_list <- list()
  
  # Iterate through each block (within each subject)
  for (block in blocks) {
    
    # Get the dataframe from the path with the subject and block name
    input_path <- file.path(base_path, "pupillometry", "raw_asc", subject, sprintf("%s_%s.asc", subject, block))
    df <- read.asc(input_path)
    
    # Get sample rate
    # PLEASE RECORD THIS IN THE LOG
    print(df$info$sample.rate)
  
    # Merge message times with sample times (from old task code split script)
    df$msg$closest_time <- sapply(df$msg$time, find_closest_time, df = df$raw)
    df$raw$closest_time <- df$raw$time # For merging
    
    combined_df <- merge(df$raw, df$msg, by = c('block', 'closest_time'), all.x = TRUE)
    combined_df$time <- combined_df$closest_time
    combined_df$subject <- subject
    combined_df <- select(combined_df, c('subject','block', 'time', 'xp', 'yp', 'ps', 'text'))
    
    
    # Adjust the 'block' column values based on block type
    if (block == 'c1') {
      combined_df$block <- combined_df$block + 40 # cond ~ 41-160
    } else if (block == 'c2') { 
      combined_df$block <- combined_df$block + 100 # 40 + 60
    } else if (block == 'g1') {
      combined_df$block <- combined_df$block + 166 # for v2, gen ~ 166-265
    } else if (block == 'g2') {
      combined_df$block <- combined_df$block + 216 # for v2, 166 + 50
    } else if (block == 'u1') {
      combined_df$block <- combined_df$block + 301 # usmem ~ 302-409
    } else if (block == 'u2') {
      combined_df$block <- combined_df$block + 355 # 301 + 54
    } else if (block == 'cm') {
      combined_df$block <- combined_df$block + 265 # csmem ~ 266-301
    } 
    
    # Store this dataframe in a list of all blocks for that subject
    block_list[[block]] <- combined_df
  }
  
  # Bind the appropriate dataframes to create 1 continuous df for cond, gen, and usmem
  cond_df <- rbind(block_list[['c1']], block_list[['c2']])
  gen_df <- rbind(block_list[['g1']], block_list[['g2']])
  usmem_df <- rbind(block_list[['u1']], block_list[['u2']])
  csmem_df <- block_list[['cm']]
  fixation_df <- block_list[['se']]
  
  print(paste0('Subject ', subject, ': Trials found in conditioning task: ', n_distinct(unique(cond_df$block))))
  print(paste0('Subject ', subject, ': Trials found in generalization task: ', n_distinct(unique(gen_df$block))))
  print(paste0('Subject ', subject, ': Trials found in us memory task: ', n_distinct(unique(usmem_df$block))))
  print(paste0('Subject ', subject, ': Trials found in cs memory task: ', n_distinct(unique(csmem_df$block))))
  
  for (phase in task_phases) {
    # Use the phase name to get the appropriate dataframe to save
    df_name <- paste0(phase, "_df")
    df_to_save <- get(df_name)
    
    # Construct the filename
    file_name <- paste0(subject, "_", phase, ".RData")
    file_path <- file.path(subject_folder, file_name)
    
    # Save
    save(df_to_save, file = file_path)
  }
  print('Completed!')
}
```

## v3

------------------------------------------------------------------------------------------------
Use this version if the subject already has interim EDFs and contains 100 trials in gen task
```{r}
# Function to find closest time
# Sometimes the time when a specific msg appears does not match with the time in raw samples, therefore we define a function to find he close time and merge raw and msg without losing any msgs
find_closest_time <- function(time, df) {
  abs_diff <- abs(df$time - time)
  closest_time <- df$time[which.min(abs_diff)]
  return(closest_time)
}

# Define subjects and base paths
subjects <- c('L031')

# Run on Mac:
base_path <- "/Users/bai/Library/CloudStorage/Box-Box/CAMSLab_Projects/Studies/AlcGen/Data/In_Lab"

asc_split_path <- file.path(base_path, "pupillometry", "restructured", "asc_split")
if (!dir.exists(asc_split_path)) {
  dir.create(asc_split_path)
}


# Create a list of the different blocks the EDFs were originally split into
blocks <- c('c1','c2','cm','g1','g2','u1','u2','se')

# Create a list of the task phases as we want to name the output files
task_phases <- c('cond','csmem','gen','usmem', 'fixation')

# Read and process data for each subject
for (subject in subjects) {
  
  # Create a folder in "asc_split" where output will go
  subject_folder <- file.path(asc_split_path, subject)
  if (!dir.exists(subject_folder)) {
    dir.create(subject_folder)
  }
  
  # Initiate an empty list to store all blocks for each subject
  block_list <- list()
  
  # Iterate through each block (within each subject)
  for (block in blocks) {
    
    # Get the dataframe from the path with the subject and block name
    input_path <- file.path(base_path, "pupillometry", "raw_asc", subject, sprintf("%s_%s.asc", subject, block))
    df <- read.asc(input_path)
    
    # Get sample rate
    # PLEASE RECORD THIS IN THE LOG
    print(df$info$sample.rate)
  
    # Merge message times with sample times (from old task code split script)
    df$msg$closest_time <- sapply(df$msg$time, find_closest_time, df = df$raw)
    df$raw$closest_time <- df$raw$time # For merging
    
    combined_df <- merge(df$raw, df$msg, by = c('block', 'closest_time'), all.x = TRUE)
    combined_df$time <- combined_df$closest_time
    combined_df$subject <- subject
    combined_df <- select(combined_df, c('subject','block', 'time', 'xp', 'yp', 'ps', 'text'))
    
    
    # Adjust the 'block' column values based on block type
    if (block == 'c1') {
      combined_df$block <- combined_df$block + 40 # cond ~ 41-160
    } else if (block == 'c2') { 
      combined_df$block <- combined_df$block + 100 # 40 + 60
    } else if (block == 'g1') {
      combined_df$block <- combined_df$block + 165 # for v3, gen ~ 166-265
    } else if (block == 'g2') {
      combined_df$block <- combined_df$block + 215 # for v3, 165 + 50
    } else if (block == 'u1') {
      combined_df$block <- combined_df$block + 301 # usmem ~ 302-409
    } else if (block == 'u2') {
      combined_df$block <- combined_df$block + 355 # 301 + 54
    } else if (block == 'cm') {
      combined_df$block <- combined_df$block + 265 # csmem ~ 266-301
    } 
    
    # Store this dataframe in a list of all blocks for that subject
    block_list[[block]] <- combined_df
  }
  
  # Bind the appropriate dataframes to create 1 continuous df for cond, gen, and usmem
  cond_df <- rbind(block_list[['c1']], block_list[['c2']])
  gen_df <- rbind(block_list[['g1']], block_list[['g2']])
  usmem_df <- rbind(block_list[['u1']], block_list[['u2']])
  csmem_df <- block_list[['cm']]
  fixation_df <- block_list[['se']]
  
  print(paste0('Subject ', subject, ': Trials found in conditioning task: ', n_distinct(unique(cond_df$block))))
  print(paste0('Subject ', subject, ': Trials found in generalization task: ', n_distinct(unique(gen_df$block))))
  print(paste0('Subject ', subject, ': Trials found in us memory task: ', n_distinct(unique(usmem_df$block))))
  print(paste0('Subject ', subject, ': Trials found in cs memory task: ', n_distinct(unique(csmem_df$block))))
  
  for (phase in task_phases) {
    # Use the phase name to get the appropriate dataframe to save
    df_name <- paste0(phase, "_df")
    df_to_save <- get(df_name)
    
    # Construct the filename
    file_name <- paste0(subject, "_", phase, ".RData")
    file_path <- file.path(subject_folder, file_name)
    
    # Save
    save(df_to_save, file = file_path)
  }
  print('Completed!')
}
```